### 2017-10-14 paper/paper.tex
Added save-generated-noise option

\documentclass{article}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{CytoGAN: Synthesis and Representation Learning of Microscopy Cell Images}

\author{
  Peter Goldsborough \\
  Facebook, Inc.\\
  Menlo Park, CA, USA \\
  \texttt{psag@fb.com} \\
  \And
  Nick Pawlowski \\
  Imperial College London \\
  London, UK \\
  \texttt{n.pawlowski16@imperial.ac.uk} \\
  \AND
  Juan C. Caicedo \\
  Imaging Platform \\
  Broad Institute of MIT and Harvard \\
  Cambridge, MA, USA \\
  \texttt{jccaicedo@broadinstitute.org} \\
  \And
  Shantanu Singh \\
  Imaging Platform \\
  Broad Institute of MIT and Harvard \\
  Cambridge, MA, USA \\
  \texttt{shsingh@broadinstitute.org} \\
  \And
  Anne E Carpenter \\
  Imaging Platform \\
  Broad Institute of MIT and Harvard \\
  Cambridge, MA, USA \\
  \texttt{anne@broadinstitute.org} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch
  (3~picas) on both the left- and right-hand margins. Use 10~point
  type, with a vertical spacing (leading) of 11~points.  The word
  \textbf{Abstract} must be centered, bold, and in point size 12. Two
  line spaces precede the abstract. The abstract must be limited to
  one paragraph.
\end{abstract}

\section{Introduction}

Advances in high-throughput microscopy systems have enabled acquisition of large volumes of high-resolution cell images, paving the way for novel computational methods that can leverage these large quantities of data, both in order to assist biologists with tasks such as morphological profiling, which aims to capture quantitative differences in the phenotypes of cellular perturbations, as well automate certain image analysis pipelines, like mechanism-of-action (MOA) prediction for drug compounds, in their entirety, end-to-end.

% The transition here feels unsmooth

Morphology-based profiling can be stated as the task of mapping microscopic cell images to representations that capture the most salient features and axes of variation of each cell's phenotype in an unsupervised manner, ideally partitioning the representation space cleanly into clusters of cells with similar visible properties and in the case of drug compounds, similar function. Two current methods that tackle this challenge successfully are specialized software like CellProfiler, which extracts representations via manually tuned features and traditional computer vision pipelines, as well as, more recently, transfer learning approaches using features learned by deep convolutional neural networks originally trained to classify miscellaneous objects.

Being an unsupervised learning problem, morphological profiling naturally invites investigation into the effectiveness of suitable approaches employing deep neural network architectures. Indeed, there have been attempts at capturing representations from cell images using autoencoders and deep variational methods, albeit with limited success. A recent addition to the unsupervised learner's toolbox is the adversarial framework introduced by Goodfellow et al. Generative adversarial networks (GANs) in particular have shown to excel at the task of synthesizing realistic natural images. In this work, we present findings on the effectiveness of GANs for two tasks in the domain of cell microscopy imaging: First, we show that GANs can generate highly realistic and biologically consistent cell images while providing greater intuition and interpretability over the synthesis mechanism than other generative methods. Second, we investigate ways to capture the internal representations of adversarial networks trained on fluorescence microscopy images of cultured human cells as a solution to morphological profiling and evaluate the quality of learned features for the task of predicting the mechanism-of-action of drug compounds.

\section{Related Work}

There exists a respectable body of work in the two lines of research that our work intersects: automated morphological profiling, and representation learning with deep neural networks, in particular generative architectures.

Caicedo et al. recently outlined the state and challenges of the morphological profiling problem in [], where the authors describe best practices for profiling pipelines using both traditional as well as deep learning methods. Prior to this, Ljosa et al. contrast the quality of representations composed of manually selected features with various augmentations and evaluate them at the task of MOA prediction on the BBBC021 dataset, consisting of microscopic images of human breast cancer cells (MCF7) treated with different compounds at differing concentrations. Both [Singh] and [Mike Ando] build on this work by showing how correcting illumination biases in images of the same dataset and applying statistical whitening transformations to learned representations, respectively, can ameliorate results reported by [Ljosa et al]. In [Nick], Pawlowski et al. for the first time report of a representation-learning method based on deep learning that is competitive with hand-picked features at the task of MOA prediction. In this work, feature layers from Inception-v3 networks trained on assorted consumer images are treated as cell representations, leveling prediction accuracy reported by [Singh].

Outside the domain of biological images, solutions to unsupervised representation learning have been proposed by [Hinton, Salakuthdinov] in the form of autoencoders and its many variations such as denoising autoencoders [], variational autoencoders [] and stacked autoencoders []. Transfer learning of convolutional network features as employed by [Nick] had first been investigated by [someone]. Preceding all of the above, Hinton et al. studied the effectiveness of restricted boltzman machines and deep belief networks at the representation learning task. More recently, the ladder network architecture was proposed by Rasmus to learn representations in a semi-supervised setting with promising results.

% Just delete this whole paragraph up there?

While both morphological profiling and unsupervised representation learning have been studied extensively in isolation, few works have synthesized the two. Nick et al. first investigated autoencoders and variational autoencoders to capture semantic representations of microscopic cell images in the BBBC021 dataset but reported results far inferior to hand-tuned features or transfer-learning approaches. The work most similar to ours is [], where Osokin et al. first study the task of cell image synthesis using generative adversarial networks and show that generated images are visually realistic and the latent space semantically smooth and biologically interpretable.

\section{Cell Image Synthesis}

In this section, we present our methodologies and findings for the task of generating realistic cell images with adversarial architectures. We begin by introducing the general GAN framework, outline the nature of the data we train on, and finally discuss our results as well as the qualitative and quantitative differences between a number of GAN models we tried.

Generative adversarial networks were first described in a seminal work by Goodfellow et al. [Foo]. The architecture they present consists of two \emph{players}, a \emph{generator} network $G$ and a \emph{discriminator} or \emph{critic} network $D$. The former receives samples $\mathbf{z} \sim P(\mathbf{z})$ drawn from a noise prior $P(z)$, typically a multi-variate uniform or normal distribution, which it maps to values $G(z)$ from the generator distribution $P_{model}$. The discriminator network must learn to distinguish such synthetic samples from real values $\mathbf{x}$ drawn from a data prior $P_{data}(\mathbf{x})$, with the goal of maximizing the distance between the generated distribution $P_{model}$ and data prior $P_{data}$. Feedback from the discriminator is in turn provided to the generator, which follows the opposite goal of fooling the discriminator into believing its synthesized samples stem from $P_{data}$. This formulation leads to a zero-sum game where the loss $L_G$ of the generator network is the negative of the discriminator's loss $L_D$, with the game's solution lying in a nash equilibrium found by playing the minimax game

$$\min_G \max_D V(G, D) = \mathbb{E}_{x \sim P_{data}}[\log(D(\mathbf{x})] + \mathbb{E}_{z \sim P_{noise}}[\log(1 - D(G(\mathbf{z}))]$$

% Trim this down ^

We evaluate the GAN framework at the task of synthesizing realistic and biologically consistent cell images. Our data prior $P_{data}$ consists of microscopy images from the BBBC021 dataset of MCF7 human breast cancer cell lines [] that were treated with 113 different drug molecules at 8 concentrations. Three \emph{channels} are available for each image in BBBC021, corresponding to DNA, F-Actin and B-Tubulin structures obtained from fluorescence microscopy. We stack these channels and treat them as RGB images via a simple $DNA \mapsto Red, Actin \mapsto Green, B-Tubulin \mapsto Blue$ mapping. Images in BBBC021 contain multiple cells and are not segmented. We segment them using the CellProfiler software, such that $P_{data}$ consists of segmented, single-cell images. Finally, we normalize the luminance of each channel to the range $[0-255]$ to effectively brighten the image, which improved the quality of generated images in our experience.

We evaluated two \emph{families} of GAN architectures. The first includes the Deep Convolutional GAN (DCGAN) model proposed by Radford et al., which models the generator network as a series of convolution and deconvolution (up-sampling?) operations that successively transform and up-sample the noise sample $z$ into an image $G(z)$ that ideally resembles those in $P_{data}$. The discriminator $D$ is a conventional deep convolutional neural network that maps images $x$ to a probability $D(x)$ representing the discriminator's confidence that the image stems from $P_{data}$. In general, the GAN game is notoriously unstable and difficult to optimize, as either player can overpower the other, leading to the discriminator giving the generator no chance to improve, or the generator exploiting only a single mode of the data distribution -- a phenomenon known as \emph{mode collapse} (this sentence needs to be better!!). We experienced this instability first-hand with the DCGAN model, with mode-collapse occurring frequently and very early on in the training process, leading to dissatisfying results. Derivations of the DCGAN such as the Least Squares (LSGAN) [cite] or Wasserstein (WGAN) [cite] architecture address the inherent instability of the GAN game and we were able to train both models to convergence. We report best qualitative results for the LSGAN model, which produced highly realistic and biologically consistent synthetic images, as shown in Figure 1.

\begin{figure}[h]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

% Need to elaborate why the images make sense biologically, like that the actin cradles the nucleus

The second family of models we successfully trained to convergence that we deem noteworthy is the class of Boundary Equilibrium GAN [cite] models proposed by [Foobar et al]. BEGAN replaces the discriminator in DCGAN with an autoencoder and uses the reconstruction loss of this autoencoder as a measure of whether the image stems from $P_{data}$ or $P_{model}$. % I think we have no room for BEGAN here?

Besides generating realistic cell images, we additionally found that interpolations between noise samples drawn from $P_{noise}$ were visually smooth. We draw two variables $z_1, z_2$ from the noise prior $P_{noise}$ and sample images for noise samples at various distances between $z_1$ and $z_2$. In all of our experiments, we used a multi-variate Gaussian distribution for the noise distribution, thus we use the spherical linear interpolation strategy from [cite] to transition between $z_1$ and $z_2$. Figure 2 shows that images change in a smooth fashion between the two points of the noise space.

\begin{figure}[h]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

Radford et al. originally reported on intriguing relationships found between regions of the noise distribution, where algebra on vectors in the noise space translated directly to corresponding semantic changes in generated images. This is similar to results studied by Mikolov et al. for word embeddings, where distances between words in the embedding space proved surprisingly rich and linear in nature. We present similar findings in Figure 3 (have to make).

% Can we add something about the biological relevance of all of this? Like why semantic relationships ^ are neat?

\section{Representation Learning}

In this section, we investigate the ability of GANs to infer vector representations of images for the purpose of profiling. We evaluate the quality of learned representations in both quantitative and qualitative experiments. First, we discuss the performance of GAN representations in the well-studied task of MOA prediction. Then, we show that representation learned by GANs have interesting semantic relationships.

The original GAN framework proposed by Goodfellow et al. does not incorporate an explicit means of inference. As such, we require extensions to traditional GAN architectures that allow us to map images drawn either from $P_{model}$ or $P_{data}$ to latent representations via some \emph{encoding} function $E(x)$. The simplest such extension is to add a hidden layer to the end of the discriminator network $D$ that induces a latent space. The intuition being that the discriminator must learn very strong internal representations in order to perform its task of separating synthetic from real images. Another benefit of this strategy is that it adds almost no computational overhead, since all but the final layer of the discriminator is shared with the encoder.  Since the BEGAN architecture models the discriminator as an autoencoder, it has a bottleneck layer that naturally provides latent representations. We compute MOA classification accuracy for LSGANs with an added hidden layer, as well as BEGANs, and report our findings in Table 1.

\begin{table}[t]
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule{1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

% Discussion?

We found that the latent representations learned by either of the two architectures did not outperform other current methods such as transfer learning or mean CellProfiler profiles. Even though MOA classification accuracy is not yet competitive, we believe that representation learning of cell images provides benefits over established methods:

\begin{itemize}
	\item Less manual tuning and faster than CellProfiler,
    \item More tuned to actual data since it is trained on cell images, providing more room for biological extensions,
    \item Results are better than for other generative models such as VAEs
\end{itemize}

\section{Conclusion}

We investigate use cases of generative adversarial networks (GANs) for the domain of cell miscroscopy imaging. First, we show that GANs have the ability to synthesize highly realistic images of cells that respect natural phenomena and explain how this is biologically useful (NOT YET). Second, we break ground on the ability of GANs to learn representations of cell images and establish their performance at the task of drug compound mechanism-of-action prediction. We discuss that even though the representations learnt by adversarial models are currently inferior at this task, GANs provide valuable benefits over each of the competing approaches.

We emphasize that this work has shed light only on the tip of the iceberg of how the adversarial framework is useful for the field of computational biology. Due to the generality of the adversarial framework, there remain many more avenues to explore. For example, we hope that future work investigates recently proposed bidirectional GAN models or other, novel approaches to infering representations under the GAN umbrella. Finally, we note that deep neural network architectures such as those described in this work flourish with large amounts of data. As such, we conjecture that the performance of approaches outlined in this work will improve if and when larger biological image datasets become available.

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
  Exploring Realistic Neural Models with the GEneral NEural SImulation
  System.}  New York: TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
learning and recall at excitatory recurrent synapses and cholinergic
modulation in rat hippocampal region CA3. {\it Journal of
  Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2017-10-14 cytogan/train/bbbc021.py

@@ -28,6 +28,7 @@ parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
+parser.add_argument('--store-generated-noise', action='store_true')
 options = common.parse_args(parser)
 
 if options.save_profiles:
@@ -334,6 +335,9 @@ with common.get_session(options.gpus) as session:
         else:
             samples = np.random.randn(options.generative_samples,
                                       model.noise_size)
+
+        if options.store_generated_noise:
+            np.savetxt('noise.csv', samples, delimiter=',')
 
         # Fix two labels and sample random noise
         if conditional_shape:

### 2017-10-15 cytogan/train/bbbc021.py
Added ability to generate images from noise

@@ -3,6 +3,7 @@
 import os
 import numpy as np
 import tensorflow as tf
+import scipy.misc
 from tqdm import tqdm
 
 from cytogan.data.cell_data import CellData
@@ -29,6 +30,7 @@ parser.add_argument('--save-profiles', action='store_true')
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
+parser.add_argument('--noise-file')
 options = common.parse_args(parser)
 
 if options.save_profiles:
@@ -337,7 +339,8 @@ with common.get_session(options.gpus) as session:
                                       model.noise_size)
 
         if options.store_generated_noise:
-            np.savetxt('noise.csv', samples, delimiter=',')
+            path = os.path.join(options.workspace, 'noise.csv')
+            np.savetxt(path, samples, delimiter=',')
 
         # Fix two labels and sample random noise
         if conditional_shape:
@@ -358,6 +361,16 @@ with common.get_session(options.gpus) as session:
                 model, [noise, labels],
                 save_to=options.figure_dir,
                 filename='generative-samples2.png')
+
+        if options.noise_file is not None:
+            noise = np.loadtxt(options.noise_file, delimiter=',')
+            images = model.generate(noise)
+            directory = os.path.join(options.workspace, 'from-noise')
+            if not os.path.exists(directory):
+                os.makedirs(directory)
+            for n, image in enumerate(images):
+                path = os.path.join(directory, '{0}.png'.format(n))
+                scipy.misc.imsave(path, image)
 
 if options.show_figures:
     visualize.show()

### 2017-10-15 cytogan/train/bbbc021.py
Update

@@ -55,13 +55,17 @@ log.debug('Options:\n%s', options.as_string)
 if not options.show_figures:
     visualize.disable_display()
 
-cell_data = CellData(options.metadata, options.labels, options.images,
-                     options.cell_count_file, options.pattern,
-                     options.normalize_luminance, options.conditional,
-                     options.concentration_only_labels)
+if not options.skip_evaluation:
+    cell_data = CellData(options.metadata, options.labels, options.images,
+                         options.cell_count_file, options.pattern,
+                         options.normalize_luminance, options.conditional,
+                         options.concentration_only_labels)
 
 image_shape = (96, 96, 3)
-number_of_batches = cell_data.number_of_images // options.batch_size
+if options.skip_training:
+    number_of_batches = 1
+else:
+    number_of_batches = cell_data.number_of_images // options.batch_size
 if options.conditional:
     # one-hot encode the compound and have a
     # continuous variable for the concentration.
@@ -364,6 +368,8 @@ with common.get_session(options.gpus) as session:
 
         if options.noise_file is not None:
             noise = np.loadtxt(options.noise_file, delimiter=',')
+            if np.ndim(noise) == 1:
+                noise = noise.reshape(1, -1)
             images = model.generate(noise)
             directory = os.path.join(options.workspace, 'from-noise')
             if not os.path.exists(directory):

### 2017-10-15 cytogan/train/common.py
Rando

@@ -129,8 +129,8 @@ def get_session(gpus):
     K.set_session(session)
     K.manual_variable_initialization(True)
 
-    np.random.seed(42)
-    tf.set_random_seed(42)
+    # np.random.seed(42)
+    # tf.set_random_seed(42)
 
     return session
 

### 2017-10-15 cytogan/train/common.py
unrandom

@@ -129,8 +129,8 @@ def get_session(gpus):
     K.set_session(session)
     K.manual_variable_initialization(True)
 
-    # np.random.seed(42)
-    # tf.set_random_seed(42)
+    np.random.seed(42)
+    tf.set_random_seed(42)
 
     return session
 

### 2017-10-15 paper/figures/algebra/make_figure.py
Adding scripts to generate figures for paper

@@ -1,0 +1,74 @@
+#!/usr/bin/env python3
+
+import argparse
+import os.path
+
+import matplotlib.pyplot as plot
+import scipy.misc
+
+parser = argparse.ArgumentParser()
+parser.add_argument('folder')
+options = parser.parse_args()
+
+
+def load(dirname, basename):
+    path = os.path.join(dirname, '{0}.png'.format(basename))
+    return scipy.misc.imread(path)
+
+
+lhs = load(options.folder, 'lhs')
+rhs = load(options.folder, 'rhs')
+base = load(options.folder, 'base')
+result = load(options.folder, 'result')
+
+figure = plot.figure(figsize=(10, 2))
+
+axis = plot.subplot(141)
+axis.get_xaxis().set_visible(False)
+axis.get_yaxis().set_visible(False)
+plot.imshow(lhs)
+
+axis.text(
+    1.22,
+    0.5,
+    r'$-$',
+    horizontalalignment='center',
+    verticalalignment='center',
+    transform=axis.transAxes,
+    fontsize=30)
+
+axis = plot.subplot(142)
+axis.get_xaxis().set_visible(False)
+axis.get_yaxis().set_visible(False)
+plot.imshow(rhs)
+
+axis.text(
+    1.22,
+    0.5,
+    r'$+$',
+    horizontalalignment='center',
+    verticalalignment='center',
+    transform=axis.transAxes,
+    fontsize=30)
+
+axis = plot.subplot(143)
+axis.get_xaxis().set_visible(False)
+axis.get_yaxis().set_visible(False)
+plot.imshow(base)
+
+axis.text(
+    1.22,
+    0.5,
+    r'$=$',
+    horizontalalignment='center',
+    verticalalignment='center',
+    transform=axis.transAxes,
+    fontsize=30)
+
+axis = plot.subplot(144)
+axis.get_xaxis().set_visible(False)
+axis.get_yaxis().set_visible(False)
+plot.imshow(result)
+
+plot.tight_layout()
+plot.savefig('figure-{0}.png'.format(options.folder))

### 2017-10-15 paper/figures/generated/make_figure.py

@@ -1,0 +1,27 @@
+#!/usr/bin/env python3
+
+import os
+
+import matplotlib.pyplot as plot
+import scipy.misc
+
+def load(path):
+    images = []
+    for file in os.listdir(path):
+        if file.endswith('.png'):
+            image = scipy.misc.imread(os.path.join(path, file))
+            images.append(image)
+    return images
+
+real = load('real')
+fake = load('fake')
+assert len(real) == len(fake), (len(real), len(fake))
+
+figure = plot.figure(figsize=(8, 2))
+for n, i in enumerate(real + fake):
+    axis = plot.subplot(2, len(real), n + 1)
+    plot.imshow(i)
+    axis.get_xaxis().set_visible(False)
+    axis.get_yaxis().set_visible(False)
+
+plot.savefig('figure.png')

### 2017-10-15 cytogan/train/visualize.py
Changed figure size for interpolations

@@ -152,7 +152,7 @@ def interpolation(model,
     if _is_grayscale(images):
         images = _make_rgb(images)
 
-    figure = plot.figure(figsize=(15, k))
+    figure = plot.figure(figsize=(interpolation_length + 2, k))
     figure.suptitle(title)
     for index, image in enumerate(images):
         _plot_image_tile(k, interpolation_length, index, image, gray)

### 2017-10-16 paper/paper2.tex
Updates to paper

\documentclass{article}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{nips_2017}
% \usepackage{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{todonotes}

%\PassOptionsToPackage{numbers}{natbib}
%\usepackage[numbers]{natbib}
%\usepackage[natbib, maxcitenames=3, style=numeric]{biblatex}
%\setcitestyle{numbers}
%\bibliographystyle{abbrvnat}

\title{Generative Modeling of Cell Images}

\author{
  Peter Goldsborough\thanks{This work has been conducted as part of an internship at the Imaging Platform of the Broad Institute.}\\
  %Imaging Platform \\
  Broad Institute of MIT and Harvard \\
  %Cambridge, MA, USA \\
  \texttt{pgoldsbo@broadinstitute.org} \\
  \And
  Nick Pawlowski \\
  %Department of Computing \\
  Imperial College London \\
  %London, UK \\
  \texttt{n.pawlowski16@imperial.ac.uk} \\
  \AND
  Juan C Caicedo \& Shantanu Singh \& Anne E Carpenter\\
  %Imaging Platform \\
  Broad Institute of MIT and Harvard \\
  %Cambridge, MA, USA \\
  \texttt{\{jccaicedo, shsingh, anne\}@broadinstitute.org} \\
}

\begin{document}

\maketitle

\begin{abstract}
We explore the application of Generative Adversarial Networks to the domain of morphological profiling of human cultured cells imaged by fluorescence microscopy. When evaluated for their ability to group cell images responding to treatment by chemicals of known classes, we find that adversarially learned representations are superior to autoencoder-based approaches. While currently inferior to classical computer vision and transfer learning, the adversarial framework enables useful and previously impossible interpretations of features due to their generative capabilities.
\end{abstract}

\section{Introduction}
Advances in high-throughput microscopy systems have enabled acquisition of large
volumes of high-resolution cell images \cite{Kraus2016}. This paves the way for
novel computational methods that can leverage these large quantities of data to
study biological systems. Our work focuses on the task of morphological
profiling, which aims to map microscopy images of cells to representation
vectors that capture salient features and axes of variation in an unsupervised
manner \cite{caicedo_profiling}. These representations ideally divide the
morphological space into clusters of cells with similar properties or, in the
case of induced perturbations, similar function.

Current techniques for morphological profiling broadly fall in two categories:
a) Classical image processing, using specialized software like CellProfiler
\cite{Carpenter2006}, to capture representations via manually-tuned segmentation
and traditional computer vision pipelines, and b) transfer learning to extract
features learned by deep convolutional neural networks originally trained to
classify miscellaneous objects \cite{pawlowski2016automating,
ando2017improving}. Classical computer vision approaches offer better
interpretability of features but require more human tuning of the segmentation
algorithms, and are limited by the feature set implemented in the image analysis
software. Current transfer learning approaches have been shown to outperform
classical methods in at least one dataset \cite{ando2017improving}. However,
given that these networks were trained on natural images (in RGB), they do not
discover the relations of the biologically meaningful image channels. Instead,
their superior performance is likely due to the their ability to capture
high-level vision features, which represent the overall structure of cells, but
not necessarily intricate details of their morphological variations. Therefore,
we hypothesized that learning representations specifically tuned to cell images
would be valuable.

We employ Generative Adversarial Networks (GANs) \cite{goodfellow2014generative}
to build a generative model of cell images from the BBBC021 dataset
\cite{Caie2010,Ljosa2012}.  We show that this generative model learns rich
feature representations and synthesizes realistic images with interpretable
properties. Our approach offers the following advantages:

\begin{itemize}
  \item \emph{Adaptability}: Unsupervised representation learning, which enables easy adaptation and incorporation of new data. This is important given that applications of morphological profiling do not have ground truth annotations, but instead aim to reveal the structure of the image collection.
  \item \emph{Specialization}: Our method learns from training data and is able to extract inherent semantic relationships. Transfer learning approaches lack this ability and cannot capture the intrinsic relationships between the biologically meaningful channels.
  \item \emph{Interpretability}: The generative abilities of this model enable previously impossible interpretbility by being able to visualize cells dependent on a given representation.
\end{itemize}

\section{Related Work}

Our work lies at the intersection of automated morphological profiling, and
representation learning with deep neural networks, in particular generative
architectures. Caicedo et al. \cite{caicedo2017data} recently outlined the state
and challenges of the morphological profiling problem. Prior to this, Ljosa et
al. \cite{Ljosa2013} compared the performance of various dimensionality
reduction methods for CellProfiler features on the BBBC021 benchmark
\cite{Caie2010,Ljosa2012}, consisting of MCF7 cells exposed to different
chemical treatments. Singh et al. \cite{Singh2014} demonstrated the importance
of illumination correction of those images. Pawlowski et al.
\cite{pawlowski2016automating} for the first time reported a
representation-learning method based on deep learning that is competitive with
hand-engineered features at the task of MOA prediction. Ando et al.
\cite{ando2017improving} also applied transfer learning with a different
architecture and introducing a novel feature normalization method. Further
related work on this dataset spans supervised classification \cite{Kraus2016a}
or deep learning applications based on CellProfiler features
\cite{Kandaswamy2016,Zamparo2015}

Few works have applied unsupervised deep learning techniques to the task of
feature extraction in morphological profiling. Pawlowski \cite{pawlowski2016msc}
first investigated autoencoder-based approaches but reported results far
inferior to hand-tuned features and transfer-learning approaches. Our model is
more related to the work by Osokin et al. \cite{osokin2017gans} and Johnson et
al. \cite{johnson2017generative}, wherein GANs were used to model cell
images, although their applications did not include morphological profiling.

\section{Using GANs for Interpretable Representation Learning}

Goodfellow et al. \cite{goodfellow2014generative} introduced GANs as a game of
two \emph{players}: a \emph{generator} $G$ and a \emph{discriminator} $D$. The
former receives samples $\mathbf{z}$ drawn from a noise prior $P(\mathbf{z}))$
which it maps to values $G(z)$ that should resemble elements of some data
distribution $P_{\text{data}}$. The discriminator must learn to distinguish such
synthetic samples from real values $\mathbf{x} \sim
P_{\text{data}}(\mathbf{x})$. The critic's confidence in the realism of the
generator's productions is used as feedback to $G$, guiding it to synthesize
ever more realistic replicates of samples from the data prior. This procedure is
formalized in a zero-sum game, solved in a nash equilibrium found by playing the
minimax game

$$\min_G \max_D V(G, D) = \mathbb{E}_{x \sim P_{data}}[\log(D(\mathbf{x})] + \mathbb{E}_{z \sim P_{noise}}[\log(1 - D(G(\mathbf{z}))].$$

Radford et al. \cite{radford2015unsupervised} first specialized GANs to image
synthesis by introducing Deep Convolutional GANs (DCGANs). DCGANs implement the
generator and discriminator based on convolutional operations. Derivations of
DCGANs such as Least Squares GAN (LSGAN) \cite{mao2016least} or Wasserstein GAN
(WGAN) \cite{arjovsky2017wasserstein} tackle infamous instabilities in the
training procedure of early generative models. In our experiments, we found
LSGAN to be most stable, in part leading to higher quality generated images than
both DCGAN and WGAN.

The original GAN framework does not include an explicit means of performing
inference. As such, we require extensions that allow mapping of a sample
$\mathbf{x}$ drawn from the data prior to a latent representation via some
encoder transformation $E(\mathbf{x})$. A common approach is to interpret the
penultimate layer of the discriminator as latent space. The activations of this
layer serve as a source of representation vectors. These latent codes are
thought to be meaningful because the discriminator must develop a strong
internal representation of its input in order to succeed at its discrimination
task. Furthermore, this method imposes no computational overhead compared to the
vanilla GAN models.

\section{Interpretable Synthesis of Cells}\label{synthesis}

GANs have been proven to synthesize realistic images both within and beyond the
biological domain
\cite{goodfellow2014generative,radford2015unsupervised,osokin2017gans}. We apply
this framework to single cell images extracted from the BBBC021 dataset.
Originally, each image consists of three \emph{channels} corresponding to DNA,
F-Actin and B-Tubulin structures obtained via fluorescence microscopy. We stack
these channels and treat them as RGB images via a simple $\text{DNA} \mapsto
\text{Red}, \text{B-Tubulin} \mapsto \text{Green}, \text{F-Actin} \mapsto
\text{Blue}$ mapping. We obtain ca. 1.3 million single cell images by segmenting
the raw images using CellProfiler. Finally, we normalize the
luminance of each channel to the range [0 -- 1].

Current approaches to morphological profiling extract features that are often
hard or impossible to understand. While CellProfiler features sometimes capture
easy concepts such as the area of the nucleus, they also features, e.g. Zernike
moments, that are difficult to interpret. For transfer learning it is nearly
impossible to visualize the captured concept. GANs present a solution as they
hold a generative model that enables the visualization of the impact of
perturbations along a feature of the noise space. Further, a more advanced
encoding method can tie the feature and noise space such that they are identical
and the visualization of feature vectors becomes possible.

\begin{figure}
  \centering
  \begin{tikzpicture}
  	\draw (0, 0) node {\includegraphics[width=0.6\textwidth]{generated}};

    \path (-3.3, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (r);
    \draw (r)+(0, -2.2) node {\textbf{Real}};

    \path (-1.07, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (l);
    \draw (l)+(0, -2.2) node {\textbf{LSGAN}};

    \path (1.14, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (w);
    \draw (w)+(0, -2.2) node {\textbf{WGAN}};

    \path (3.35, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (d);
    \draw (d)+(0, -2.2) node {\textbf{DCGAN}};
  \end{tikzpicture}
  \caption{Examples of real MCF7 cells from the BBBC021 dataset juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN. The generated images of LSGAN are most realistic and consistent with biological properties of cells.}
  \label{fig:generated}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth, trim={0 0 0 2.5cm}, clip]{interpolation}
  \caption{Interpolating between two points $\mathbf{z_1}, \mathbf{z_2} \sim P(\mathbf{z}))$ results in visually smooth transitions in the corresponding synthesized images. Each row shows the transition from $\mathbf{z_1}$ to $\mathbf{z_2}$ from left to right.}
  \label{fig:interpolations}
  \vspace{-0.05cm}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{algebra-tubulin}
  \includegraphics[width=0.6\textwidth]{algebra-nucleus}
  \caption{Vector algebra in the noise space translates to biologically valuable relationships in generated images. In the first row, subtracting the vector for a cell with small quantities of B-Tubulin (green stain) from one with high amounts yields a vector representation of "high B-Tubulin content". In the second row, the difference between a large and small nucleus encodes the semantic meaning of "larger nucleus", which can be added to vectors of other cell images to grow the size of the nucleus.}
  \label{fig:algebra}
\end{figure}

We show the generative abilities of our models and the richness of the noise space on three ways: a) Generation of high quality cell images as shown in Figure \ref{fig:generated}. The LSGAN produces the most convincing images. b) The noise space is capable of smooth interpolations between two noise samples $\mathbf{z_1}, \mathbf{z_2} \sim P(\mathbf{z}))$, displayed in Figure \ref{fig:interpolations}. This undermines the fact that the generator learned low-dimensional manifold of the images. c) $P(\mathbf{z}))$ encodes semantic information that enables algebra based on semantically relevant properties of the images. Figure \ref{fig:algebra} presents examples of vector algebra in $P(\mathbf{z})$ with biologically interpretable results. We are able to leverage the linear relationships present in the latent space to influence the size of synthesized nuclei as well as B-Tubulin content.

\section{Representation Learning for Morphological Profiling}\label{moa}
We test the quality of representations extracted by the discriminator used to generate single cell images learned by testing their ability to cluster treatments of similar function or mechanism-of-action. We obtain treatment profiles by averaging the extracted single cell representations found as intermediate activations of the discriminator. Further, we follow the experimental protocol of \cite{Ljosa2013} and report an average mechanism-of-action classification accuracy of a leave-one-compound-out cross-validation using a one nearest neighbor classifier.

\begin{table}
  \centering
  \begin{tabular}{lllllll}
  	\toprule
    DCGAN & WGAN & LSGAN & VAE \cite{pawlowski2016msc} & CP \cite{Singh2014} & Deep Transfer \cite{pawlowski2016automating}  & Deep Transfer \cite{ando2017improving} \\
    \midrule
    43 \%& 45 \% & 63 \% & 49 \% & 90 \% & 91 \% & 96 \% \\
    \bottomrule \\
  \end{tabular}
  \caption{Accuracy of various methods at the task of predicting the mechanism-of-action of treatments via nearest-neighbor classification in the BBBC021 dataset. CP refers to the results by Singh et al. \cite{Singh2014} using CellProfiler Features. The Deep Transfer methods correspond to Deep Feature Transfer as proposed by Pawlowski et al. \cite{pawlowski2016automating} and refined by Ando et al. \cite{ando2017improving}.}
  \label{moa-results}
  \vspace{-0.1cm}
\end{table}

Table \ref{moa-results} compares the accuracy of our approach to classical CellProfiler features and transfer learning. We found that the quality of representations found by unsupervised learning is not yet competitive with other established methods. Nevertheless, we believe that the adversarial approach has significant benefits regarding its interpretability as outlined above. Additionally, this framework is able to adapt to the dataset at hand and extract inherent relations that are not captured by previous methods. We believe that further gains can be made, as the image quality corresponds to classification accuracy. As such the best performing method (LSGAN) also yields the highest image quality, as judged qualitatively by expert biologists and shown in Figure \ref{fig:generated}.

\section{Conclusion}
This work investigates the use of Generative Adversarial Networks for the domain
of cell microscopy imaging, particularly morphological profiling. First, we
demonstrate that GANs are able to synthesize realistic and biologically
consistent images of cells. In addition, we explain that GANs induce a rich
semantic structure in their noise space, providing a powerful basis for
interpretation of observed properties. Second, we test the quality of the
representations learned via an encoding extension to standard GANs by evaluating
their mechanism-of-action classification performance following \cite{Ljosa2013}.
Even though adversarially learned representations are currently inferior at this
task, we argue that further enhancements to GANs, especially biologically
meaningful ones, will lead to richer latent representations. We base this claim
on our observation that better generative models exhibit better classification
accuracies at this task.

We emphasize that this work covers only a very small fraction of possible
applications of GANs to the domain of microscopical images. For example, we hope
that future work investigates BiGANs \cite{donahue2016adversarial} or other
novel solutions to infer latent features with GANs. We believe that improved
inference, combined with the interpretable nature of the GAN framework, may
enable simulated experiments by performing algebra with vectors corresponding to
cell lines, diseases or other perturbations. Finally, we note that deep learning
generally flourishes with larger amounts of data. As such, we conjecture that
the performance of the outlined approaches will improve with the use of larger
datasets.

\section*{Acknowledgements}
Nick Pawlowski is supported by Microsoft Research through its PhD Scholarship
Program and the EPSRC Centre for Doctoral Training in High Performance
Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).

\bibliographystyle{plain}
\bibliography{lit}

\end{document}

### 2017-10-16 cytogan/train/bbbc021.py

@@ -301,6 +301,8 @@ with common.get_session(options.gpus) as session:
     if options.interpolate_samples is not None:
         start, end = np.random.randn(2, options.interpolate_samples[0],
                                      model.noise_size)
+        start = start[[7, 9, 13], :]
+        end = end[[7, 9, 13], :]
         if conditional_shape:
             labels = cell_data.sample_labels(options.interpolate_samples[0])
             labels = np.array(list(labels))
@@ -311,7 +313,8 @@ with common.get_session(options.gpus) as session:
             model,
             start,
             end,
-            *options.interpolate_samples,
+            len(start),
+            options.interpolate_samples[1],
             options.interpolation_method,
             conditional=labels,
             gray=True,

### 2017-10-16 cytogan/train/common.py

@@ -129,7 +129,7 @@ def get_session(gpus):
     K.set_session(session)
     K.manual_variable_initialization(True)
 
-    np.random.seed(42)
+    np.random.seed(16)
     tf.set_random_seed(42)
 
     return session

### 2017-10-16 cytogan/train/visualize.py

@@ -33,7 +33,7 @@ def _save_figure(folder, filename):
     if not os.path.exists(folder):
         os.makedirs(folder)
     log.info('Saving %s', path)
-    plot.savefig(path)
+    plot.savefig(path, bbox_inches='tight', pad_inches=0)
 
 
 def reconstructions(model,
@@ -152,10 +152,13 @@ def interpolation(model,
     if _is_grayscale(images):
         images = _make_rgb(images)
 
-    figure = plot.figure(figsize=(interpolation_length + 2, k))
-    figure.suptitle(title)
+    figure = plot.figure(figsize=(8, k))
+    # figure.suptitle(title)
     for index, image in enumerate(images):
         _plot_image_tile(k, interpolation_length, index, image, gray)
+
+    plot.subplots_adjust(left=0.1, right=0.9, top=0.6, bottom=0.2)
+    plot.tight_layout()
 
     if save_to is not None:
         filename = '{0}-interpolation.png'.format(method)

### 2017-10-16 paper/figures/generated/make_figure.py

@@ -14,14 +14,20 @@ def load(path):
     return images
 
 real = load('real')
-fake = load('fake')
-assert len(real) == len(fake), (len(real), len(fake))
+lsgan = load('lsgan')
+wgan = load('wgan')
+dcgan = load('dcgan')
 
-figure = plot.figure(figsize=(8, 2))
-for n, i in enumerate(real + fake):
-    axis = plot.subplot(2, len(real), n + 1)
+images = []
+for i in range(2):
+    images += [x[i] for x in (real, lsgan, wgan, dcgan)]
+
+figure = plot.figure(figsize=(5, 2))
+for n, i in enumerate(images):
+    axis = plot.subplot(2, 4, n + 1)
+    axis.axis('off')
     plot.imshow(i)
     axis.get_xaxis().set_visible(False)
     axis.get_yaxis().set_visible(False)
 
-plot.savefig('figure.png')
+plot.savefig('figure.png', bbox_inches='tight', pad_inches=0)

### 2017-10-16 paper/paper.tex
Update to paper

@@
 \usepackage[final]{nips_2017}
+% \usepackage{nips_2017}
 
@@
 \usepackage{booktabs}       % professional-quality tables
+\usepackage{amsmath}
 \usepackage{amsfonts}       % blackboard math symbols
@@
 \usepackage{microtype}      % microtypography
+\usepackage{todonotes}
 
-\title{CytoGAN: Synthesis and Representation Learning of Microscopy Cell Images}
+%\PassOptionsToPackage{numbers}{natbib}
+%\usepackage[numbers]{natbib}
+%\usepackage[natbib, maxcitenames=3, style=numeric]{biblatex}
+%\setcitestyle{numbers}
+%\bibliographystyle{abbrvnat}
+
+\title{Generative Modeling of Cell Images}
 
 \author{
-  Peter Goldsborough \\
-  Facebook, Inc.\\
-  Menlo Park, CA, USA \\
-  \texttt{psag@fb.com} \\
+  Peter Goldsborough\thanks{This work has been conducted as part of an internship at the Imaging Platform of the Broad Institute.}\\
+  %Imaging Platform \\
+  Broad Institute of MIT and Harvard \\
+  %Cambridge, MA, USA \\
+  \texttt{pgoldsbo@broadinstitute.org} \\
   \And
   Nick Pawlowski \\
+  %Department of Computing \\
   Imperial College London \\
-  London, UK \\
+  %London, UK \\
   \texttt{n.pawlowski16@imperial.ac.uk} \\
   \AND
-  Juan C. Caicedo \\
-  Imaging Platform \\
+  Juan C Caicedo \& Shantanu Singh \& Anne E Carpenter\\
+  %Imaging Platform \\
   Broad Institute of MIT and Harvard \\
-  Cambridge, MA, USA \\
-  \texttt{jccaicedo@broadinstitute.org} \\
-  \And
-  Shantanu Singh \\
-  Imaging Platform \\
-  Broad Institute of MIT and Harvard \\
-  Cambridge, MA, USA \\
-  \texttt{shsingh@broadinstitute.org} \\
-  \And
-  Anne E Carpenter \\
-  Imaging Platform \\
-  Broad Institute of MIT and Harvard \\
-  Cambridge, MA, USA \\
-  \texttt{anne@broadinstitute.org} \\
+  %Cambridge, MA, USA \\
+  \texttt{\{jccaicedo, shsingh, anne\}@broadinstitute.org} \\
 }
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch
-  (3~picas) on both the left- and right-hand margins. Use 10~point
-  type, with a vertical spacing (leading) of 11~points.  The word
-  \textbf{Abstract} must be centered, bold, and in point size 12. Two
-  line spaces precede the abstract. The abstract must be limited to
-  one paragraph.
+We explore the application of Generative Adversarial Networks to the domain of morphological profiling of human cultured cells imaged by fluorescence microscopy.
+When evaluated for their ability to group cell images responding to treatment by chemicals of known classes, we find that adversarially learned representations are superior to autoencoder-based approaches.
+While currently inferior to classical computer vision and transfer learning, the adversarial framework enables useful and previously impossible interpretability of latent codes.
 \end{abstract}
@@
 \section{Introduction}
+Advances in high-throughput microscopy systems have enabled acquisition of large
+volumes of high-resolution cell images \cite{Kraus2016}. This paves the way for
+novel computational methods that can leverage these large quantities of data to
+study biological systems. Our work focuses on the task of morphological
+profiling, which aims to map microscopy images of cells to representation
+vectors that capture salient features and axes of variation in an unsupervised
+manner \cite{caicedo_profiling}. These representations ideally divide the
+morphological space into clusters of cells with similar properties or, in the
+case of induced perturbations, similar function.
 
-Advances in high-throughput microscopy systems have enabled acquisition of large volumes of high-resolution cell images, paving the way for novel computational methods that can leverage these large quantities of data, both in order to assist biologists with tasks such as morphological profiling, which aims to capture quantitative differences in the phenotypes of cellular perturbations, as well automate certain image analysis pipelines, like mechanism-of-action (MOA) prediction for drug compounds, in their entirety, end-to-end.
+Current techniques for morphological profiling broadly fall in two categories:
+a) Classical image processing, using specialized software like CellProfiler
+\cite{Carpenter2006}, to capture representations via manually-tuned segmentation
+and traditional computer vision pipelines, and b) transfer learning to extract
+features learned by deep convolutional neural networks originally trained to
+classify miscellaneous objects \cite{pawlowski2016automating,
+ando2017improving}. Classical computer vision approaches offer better
+interpretability of features but require more human tuning of the segmentation
+algorithms, and are limited by the feature set implemented in the image analysis
+software. Current transfer learning approaches have been shown to outperform
+classical methods in at least one dataset \cite{ando2017improving}. However,
+given that these networks were trained on natural images (in RGB), they do not
+discover the relations of the biologically meaningful image channels. Instead,
+their superior performance is likely due to the their ability to capture
+high-level vision features, which represent the overall structure of cells, but
+not necessarily intricate details of their morphological variations. Therefore,
+we hypothesized that learning representations specifically tuned to cell images
+would be valuable.
 
-% The transition here feels unsmooth
+We employ Generative Adversarial Networks (GANs) \cite{goodfellow2014generative}
+to build a generative model of cell images from the BBBC021 dataset
+\cite{Caie2010,Ljosa2012}.  We show that this generative model learns rich
+feature representations and synthesizes realistic images with interpretable
+properties. Our approach offers the following advantages:
 
-Morphology-based profiling can be stated as the task of mapping microscopic cell images to representations that capture the most salient features and axes of variation of each cell's phenotype in an unsupervised manner, ideally partitioning the representation space cleanly into clusters of cells with similar visible properties and in the case of drug compounds, similar function.
-Two current methods that tackle this challenge successfully are specialized software like CellProfiler, which extracts representations via manually tuned features and traditional computer vision pipelines, as well as, more recently, transfer learning approaches using features learned by deep convolutional neural networks originally trained to classify miscellaneous objects.
-
-Being an unsupervised learning problem, morphological profiling naturally invites investigation into the effectiveness of suitable approaches employing deep neural network architectures.
-Indeed, there have been attempts at capturing representations from cell images using autoencoders and deep variational methods, albeit with limited success.
-A recent addition to the unsupervised learner's toolbox is the adversarial framework introduced by Goodfellow et al.
-Generative adversarial networks (GANs) in particular have shown to excel at the task of synthesizing realistic natural images.
-In this work, we present findings on the effectiveness of GANs for two tasks in the domain of cell microscopy imaging: First, we show that GANs can generate highly realistic and biologically consistent cell images while providing greater intuition and interpretability over the synthesis mechanism than other generative methods.
-Second, we investigate ways to capture the internal representations of adversarial networks trained on fluorescence microscopy images of cultured human cells as a solution to morphological profiling and evaluate the quality of learned features for the task of predicting the mechanism-of-action of drug compounds.
+\begin{itemize}
+  \item \emph{Adaptability}: Unsupervised representation learning, which enables easy adaptation and incorporation of new data.
+This is important given that applications of morphological profiling do not have ground truth annotations, but instead aim to reveal the structure of the image collection.
+  \item \emph{Specialization}: Our method learns from training data and is able to extract inherent semantic relationships.
+Transfer learning approaches lack this ability and cannot capture the intrinsic relationships between the biologically meaningful channels.
+  \item \emph{Interpretability}: The generative abilities of this model enable previously impossible interpretbility by being able to visualize cells dependent on a given representation.
+\end{itemize}
 
@@
 
-There exists a respectable body of work in the two lines of research that our work intersects: automated morphological profiling, and representation learning with deep neural networks, in particular generative architectures.
+Our work lies at the intersection of automated morphological profiling, and
+representation learning with deep neural networks, in particular generative
+architectures. Caicedo et al. \cite{caicedo2017data} recently outlined the state
+and challenges of the morphological profiling problem. Prior to this, Ljosa et
+al. \cite{Ljosa2013} compared the performance of various dimensionality
+reduction methods for CellProfiler features on the BBBC021 benchmark
+\cite{Caie2010,Ljosa2012}, consisting of MCF7 cells exposed to different
+chemical treatments. Singh et al. \cite{Singh2014} demonstrated the importance
+of illumination correction of those images. Pawlowski et al.
+\cite{pawlowski2016automating} for the first time reported a
+representation-learning method based on deep learning that is competitive with
+hand-engineered features at the task of MOA prediction. Ando et al.
+\cite{ando2017improving} also applied transfer learning with a different
+architecture and introducing a novel feature normalization method. Further
+related work on this dataset spans supervised classification \cite{Kraus2016a}
+or deep learning applications based on CellProfiler features
+\cite{Kandaswamy2016,Zamparo2015}
 
-Caicedo et al.
-recently outlined the state and challenges of the morphological profiling problem in [], where the authors describe best practices for profiling pipelines using both traditional as well as deep learning methods.
-Prior to this, Ljosa et al.
-contrast the quality of representations composed of manually selected features with various augmentations and evaluate them at the task of MOA prediction on the BBBC021 dataset, consisting of microscopic images of human breast cancer cells (MCF7) treated with different compounds at differing concentrations.
-Both [Singh] and [Mike Ando] build on this work by showing how correcting illumination biases in images of the same dataset and applying statistical whitening transformations to learned representations, respectively, can ameliorate results reported by [Ljosa et al].
-In [Nick], Pawlowski et al.
-for the first time report of a representation-learning method based on deep learning that is competitive with hand-picked features at the task of MOA prediction.
-In this work, feature layers from Inception-v3 networks trained on assorted consumer images are treated as cell representations, leveling prediction accuracy reported by [Singh].
+Few works have applied unsupervised deep learning techniques to the task of
+feature extraction in morphological profiling. Pawlowski \cite{pawlowski2016msc}
+first investigated autoencoder-based approaches but reported results far
+inferior to hand-tuned features and transfer-learning approaches. Our model is
+more related to the work by Osokin et al. \cite{osokin2017gans} and Johnson et
+al. \cite{johnson2017generative}, wherein GANs were used to model cell
+images, although their applications did not include morphological profiling.
 
-Outside the domain of biological images, solutions to unsupervised representation learning have been proposed by [Hinton, Salakuthdinov] in the form of autoencoders and its many variations such as denoising autoencoders [], variational autoencoders [] and stacked autoencoders [].
-Transfer learning of convolutional network features as employed by [Nick] had first been investigated by [someone].
-Preceding all of the above, Hinton et al.
-studied the effectiveness of restricted boltzman machines and deep belief networks at the representation learning task.
-More recently, the ladder network architecture was proposed by Rasmus to learn representations in a semi-supervised setting with promising results.
+\section{Using GANs for Interpretable Representation Learning}
 
-% Just delete this whole paragraph up there?
+Goodfellow et al. \cite{goodfellow2014generative} introduced GANs as a game of
+two \emph{players}: a \emph{generator} $G$ and a \emph{discriminator} $D$. The
+former receives samples $\mathbf{z}$ drawn from a noise prior $P(\mathbf{z})$
+which it maps to values $G(z)$ that should resemble elements of some data
+distribution $P_{\text{data}}$. The discriminator must learn to distinguish such
+synthetic samples from real values $\mathbf{x} \sim
+P_{\text{data}}(\mathbf{x})$. The critic's confidence in the realism of the
+generator's productions is used as feedback to $G$, guiding it to synthesize
+ever more realistic replicates of samples from the data prior. This procedure is
+formalized in a zero-sum game, solved in a nash equilibrium found by playing the
+minimax game
 
-While both morphological profiling and unsupervised representation learning have been studied extensively in isolation, few works have synthesized the two.
-Nick et al.
-first investigated autoencoders and variational autoencoders to capture semantic representations of microscopic cell images in the BBBC021 dataset but reported results far inferior to hand-tuned features or transfer-learning approaches.
-The work most similar to ours is [], where Osokin et al.
-first study the task of cell image synthesis using generative adversarial networks and show that generated images are visually realistic and the latent space semantically smooth and biologically interpretable.
+$$\min_G \max_D V(G, D) = \mathbb{E}_{x \sim P_{data}}[\log(D(\mathbf{x})] + \mathbb{E}_{z \sim P_{noise}}[\log(1 - D(G(\mathbf{z}))].$$
 
-\section{Cell Image Synthesis}
+Radford et al. \cite{radford2015unsupervised} first specialized GANs to image
+synthesis by introducing Deep Convolutional GANs (DCGANs). DCGANs implement the
+generator and discriminator based on convolutional operations. Derivations of
+DCGANs such as Least Squares GAN (LSGAN) \cite{mao2016least} or Wasserstein GAN
+(WGAN) \cite{arjovsky2017wasserstein} tackle infamous instabilities in the
+training procedure of early generative models. In our experiments, we found
+LSGAN to be most stable, in part leading to higher quality generated images than
+both DCGAN and WGAN.
 
-In this section, we present our methodologies and findings for the task of generating realistic cell images with adversarial architectures.
-We begin by introducing the general GAN framework, outline the nature of the data we train on, and finally discuss our results as well as the qualitative and quantitative differences between a number of GAN models we tried.
+The original GAN framework does not include an explicit means of performing
+inference. As such, we require extensions that allow mapping of a sample
+$\mathbf{x}$ drawn from the data prior to a latent representation via some
+encoder transformation $E(\mathbf{x})$. A common approach is to interpret the
+penultimate layer of the discriminator as latent space. The activations of this
+layer serve as a source of representation vectors. These latent codes are
+thought to be meaningful because the discriminator must develop a strong
+internal representation of its input in order to succeed at its discrimination
+task. Furthermore, this method imposes no computational overhead compared to the
+vanilla GAN models.
 
-Generative adversarial networks were first described in a seminal work by Goodfellow et al.
-[Foo].
-The architecture they present consists of two \emph{players}, a \emph{generator} network $G$ and a \emph{discriminator} or \emph{critic} network $D$.
-The former receives samples $\mathbf{z} \sim P(\mathbf{z})$ drawn from a noise prior $P(z)$, typically a multi-variate uniform or normal distribution, which it maps to values $G(z)$ from the generator distribution $P_{model}$.
-The discriminator network must learn to distinguish such synthetic samples from real values $\mathbf{x}$ drawn from a data prior $P_{data}(\mathbf{x})$, with the goal of maximizing the distance between the generated distribution $P_{model}$ and data prior $P_{data}$.
-Feedback from the discriminator is in turn provided to the generator, which follows the opposite goal of fooling the discriminator into believing its synthesized samples stem from $P_{data}$.
-This formulation leads to a zero-sum game where the loss $L_G$ of the generator network is the negative of the discriminator's loss $L_D$, with the game's solution lying in a nash equilibrium found by playing the minimax game
+\section{Interpretable Synthesis of Cells}\label{synthesis}
 
-$$\min_G \max_D V(G, D) = \mathbb{E}_{x \sim P_{data}}[\log(D(\mathbf{x})] + \mathbb{E}_{z \sim P_{noise}}[\log(1 - D(G(\mathbf{z}))]$$
+We examine if GAN models' reported ability to synthesize realistic images
+transfers to the domain of cell microscopy. Our data prior consists of images
+from the BBBC021 dataset of human breast cancer cells. Originally, each image
+consists of three \emph{channels} corresponding to DNA, F-Actin and B-Tubulin,
+cellular components visualized by fluorescence microscopy. We stack these
+channels and treat them as RGB images via a simple $\text{DNA} \mapsto \text{R},
+\text{$\beta$-Tubulin} \mapsto \text{G}, \text{F-Actin} \mapsto \text{B}$
+mapping. We obtain ca. 1.3 million single cell images by segmenting the raw
+images using CellProfiler software. Finally, we normalize the luminance of each
+channel to the range [0 -- 1].
 
-% Trim this down ^
+Figure \ref{fig:generated} shows examples of images generated with LSGAN, WGAN
+and DCGAN architectures alongside real images. The synthetic images,
+particularly those produced by LSGAN, are not only highly detailed and
+realistic, but also consistent with their biological nature. For example, it is
+characteristic that B-Tubulin forms a circular halo cradling the nucleus of a
+cell. This property is maintained clearly in all generated images.
 
-We evaluate the GAN framework at the task of synthesizing realistic and biologically consistent cell images.
-Our data prior $P_{data}$ consists of microscopy images from the BBBC021 dataset of MCF7 human breast cancer cell lines [] that were treated with 113 different drug molecules at 8 concentrations.
-Three \emph{channels} are available for each image in BBBC021, corresponding to DNA, F-Actin and B-Tubulin structures obtained from fluorescence microscopy.
-We stack these channels and treat them as RGB images via a simple $DNA \mapsto Red, Actin \mapsto Green, B-Tubulin \mapsto Blue$ mapping.
-Images in BBBC021 contain multiple cells and are not segmented.
-We segment them using the CellProfiler software, such that $P_{data}$ consists of segmented, single-cell images.
-Finally, we normalize the luminance of each channel to the range $[0-255]$ to effectively brighten the image, which improved the quality of generated images in our experience.
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \tikzset{frame/.style={draw, semithick, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt}}
 
-We evaluated two \emph{families} of GAN architectures.
-The first includes the Deep Convolutional GAN (DCGAN) model proposed by Radford et al., which models the generator network as a series of convolution and deconvolution (up-sampling?) operations that successively transform and up-sample the noise sample $z$ into an image $G(z)$ that ideally resembles those in $P_{data}$.
-The discriminator $D$ is a conventional deep convolutional neural network that maps images $x$ to a probability $D(x)$ representing the discriminator's confidence that the image stems from $P_{data}$.
-In general, the GAN game is notoriously unstable and difficult to optimize, as either player can overpower the other, leading to the discriminator giving the generator no chance to improve, or the generator exploiting only a single mode of the data distribution -- a phenomenon known as \emph{mode collapse} (this sentence needs to be better!!).
-We experienced this instability first-hand with the DCGAN model, with mode-collapse occurring frequently and very early on in the training process, leading to dissatisfying results.
-Derivations of the DCGAN such as the Least Squares (LSGAN) [cite] or Wasserstein (WGAN) [cite] architecture address the inherent instability of the GAN game and we were able to train both models to convergence.
-We report best qualitative results for the LSGAN model, which produced highly realistic and biologically consistent synthetic images, as shown in Figure 1.
+  	\draw (0, 0) node {\includegraphics[width=0.6\textwidth]{generated}};
 
-\begin{figure}[h]
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
+    \path (-3.3, 0) coordinate [frame] (r);
+    \draw (r)+(0, -2.2) node {\textbf{Real}};
+
+    \path (-1.07, 0) coordinate [frame] (l);
+    \draw (l)+(0, -2.2) node {\textbf{LSGAN}};
+
+    \path (1.14, 0) coordinate [frame] (w);
+    \draw (w)+(0, -2.2) node {\textbf{WGAN}};
+
+    \path (3.35, 0) coordinate [frame] (d);
+    \draw (d)+(0, -2.2) node {\textbf{DCGAN}};
+  \end{tikzpicture}
+  \caption{Examples of real MCF7 cells from the BBBC021 dataset juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN.
+The generated images of LSGAN are most realistic and consistent with biological properties of cells.}
+  \label{fig:generated}
 \end{figure}
 
-% Need to elaborate why the images make sense biologically, like that the actin cradles the nucleus
+Current approaches to morphological profiling capture representations that are
+often hard to interpret. While certain CellProfiler features such as nucleic
+area are easy to understand, others like Zernike moments are not intuitive. For
+transfer learning it is nearly impossible to visualize the captured concept. In
+contrast, the GAN noise space $P(z)$ has been shown to be highly interpretable
+and reveal rich semantic structure \cite{radford2015unsupervised}. We are able
+to demonstrate this for images of cells. Figure \ref{fig:interpolations}
+exhibits how interpolating between two noise samples $z_1, z_2 \sim P(z)$ leads
+to visually smooth transitions in synthesized cells. Figure \ref{fig:algebra}
+shows that $P(z)$ encodes semantic relationships, enabling algebra on
+interpretable properties of the images. While there is no way to encode images
+into $P(\mathbf{z})$ with the methods presented so far, we believe more advanced
+models that enable this will be highly valuable if they can maintain these rich
+semantic properties.
 
-The second family of models we successfully trained to convergence that we deem noteworthy is the class of Boundary Equilibrium GAN [cite] models proposed by [Foobar et al].
-BEGAN replaces the discriminator in DCGAN with an autoencoder and uses the reconstruction loss of this autoencoder as a measure of whether the image stems from $P_{data}$ or $P_{model}$.
-% I think we have no room for BEGAN here?
-
-Besides generating realistic cell images, we additionally found that interpolations between noise samples drawn from $P_{noise}$ were visually smooth.
-We draw two variables $z_1, z_2$ from the noise prior $P_{noise}$ and sample images for noise samples at various distances between $z_1$ and $z_2$.
-In all of our experiments, we used a multi-variate Gaussian distribution for the noise distribution, thus we use the spherical linear interpolation strategy from [cite] to transition between $z_1$ and $z_2$.
-Figure 2 shows that images change in a smooth fashion between the two points of the noise space.
-
-\begin{figure}[h]
+\begin{figure}
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
+  \includegraphics[width=\textwidth, trim={0 0 0 2.5cm}, clip]{interpolation}
+  \caption{Interpolating between two points $\mathbf{z_1}, \mathbf{z_2} \sim P(\mathbf{z})$ results in visually smooth transitions in the corresponding synthesized images.
+Each row shows the transition from $\mathbf{z_1}$ to $\mathbf{z_2}$ from left to right.}
+  \vspace{-0.1cm}
 \end{figure}
 
-Radford et al.
-originally reported on intriguing relationships found between regions of the noise distribution, where algebra on vectors in the noise space translated directly to corresponding semantic changes in generated images.
-This is similar to results studied by Mikolov et al.
-for word embeddings, where distances between words in the embedding space proved surprisingly rich and linear in nature.
-We present similar findings in Figure 3 (have to make).
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\textwidth]{algebra-tubulin}
+  \includegraphics[width=0.6\textwidth]{algebra-nucleus}
+  \caption{Vector algebra in the noise space translates to biologically valuable relationships in generated images.
+In the first row, subtracting the vector for a cell with small quantities of $\beta$-Tubulin (green stain) from one with high amounts yields a vector representation for "high $\beta$-Tubulin content".
+In the second row, the difference between a large and small nucleus encodes the semantic meaning of "larger nucleus", which can be added to vectors of other cell images to grow the size of the nucleus.}
+  \label{fig:algebra}
+\end{figure}
 
-% Can we add something about the biological relevance of all of this? Like why semantic relationships ^ are neat?
+\section{Representation Learning for Morphological Profiling}\label{moa}
+We test the quality of representations extracted by the discriminator used to generate single cell images learned by testing their ability to cluster treatments of similar function or mechanism-of-action.
+We obtain treatment profiles by averaging the extracted single cell representations found as intermediate activations of the discriminator.
+Further, we follow the experimental protocol of \cite{Ljosa2013} and report an average mechanism-of-action classification accuracy of a leave-one-compound-out cross-validation using a one nearest neighbor classifier.
 
-\section{Representation Learning}
-
-In this section, we investigate the ability of GANs to infer vector representations of images for the purpose of profiling.
-We evaluate the quality of learned representations in both quantitative and qualitative experiments.
-First, we discuss the performance of GAN representations in the well-studied task of MOA prediction.
-Then, we show that representation learned by GANs have interesting semantic relationships.
-
-The original GAN framework proposed by Goodfellow et al.
-does not incorporate an explicit means of inference.
-As such, we require extensions to traditional GAN architectures that allow us to map images drawn either from $P_{model}$ or $P_{data}$ to latent representations via some \emph{encoding} function $E(x)$.
-The simplest such extension is to add a hidden layer to the end of the discriminator network $D$ that induces a latent space.
-The intuition being that the discriminator must learn very strong internal representations in order to perform its task of separating synthetic from real images.
-Another benefit of this strategy is that it adds almost no computational overhead, since all but the final layer of the discriminator is shared with the encoder.
-Since the BEGAN architecture models the discriminator as an autoencoder, it has a bottleneck layer that naturally provides latent representations.
-We compute MOA classification accuracy for LSGANs with an added hidden layer, as well as BEGANs, and report our findings in Table 1.
-
-\begin{table}[t]
-  \caption{Sample table title}
-  \label{sample-table}
+\begin{table}
   \centering
-  \begin{tabular}{lll}
+  \begin{tabular}{lllllll}
   	\toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule{1-2}
-    Name     & Description     & Size ($\mu$m) \\
+    DCGAN & WGAN & LSGAN & VAE \cite{pawlowski2016msc} & CP \cite{Singh2014} & Deep Transfer \cite{pawlowski2016automating}  & Deep Transfer \cite{ando2017improving} \\
     \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
+    43 \%& 45 \% & 63 \% & 49 \% & 90 \% & 91 \% & 96 \% \\
+    \bottomrule \\
   \end{tabular}
+  \caption{Accuracy of various methods at the task of predicting the mechanism-of-action of treatments via nearest-neighbor classification in the BBBC021 dataset.
+CP refers to the results by Singh et al.
+\cite{Singh2014} using CellProfiler Features.
+The Deep Transfer methods correspond to Deep Feature Transfer as proposed by Pawlowski et al.
+\cite{pawlowski2016automating} and refined by Ando et al.
+\cite{ando2017improving}.}
+  \label{moa-results}
 \end{table}
 
-% Discussion?
-
-We found that the latent representations learned by either of the two architectures did not outperform other current methods such as transfer learning or mean CellProfiler profiles.
-Even though MOA classification accuracy is not yet competitive, we believe that representation learning of cell images provides benefits over established methods:
-
-\begin{itemize}
-	\item Less manual tuning and faster than CellProfiler,
-    \item More tuned to actual data since it is trained on cell images, providing more room for biological extensions,
-    \item Results are better than for other generative models such as VAEs
-\end{itemize}
+Table \ref{moa-results} compares the accuracy of our approach to classical CellProfiler features and transfer learning.
+We found that the quality of representations found by unsupervised learning is not yet competitive with other established methods.
+Nevertheless, we believe that the adversarial approach has significant benefits regarding its interpretability as outlined above.
+Additionally, this framework is able to adapt to the dataset at hand and extract inherent relations that are not captured by previous methods.
+We believe that further gains can be made, as the image quality corresponds to classification accuracy.
+As such the best performing method (LSGAN) also yields the highest image quality, as judged qualitatively by expert biologists and shown in Figure \ref{fig:generated}.
 
@@
 
-We investigate use cases of generative adversarial networks (GANs) for the domain of cell miscroscopy imaging.
-First, we show that GANs have the ability to synthesize highly realistic images of cells that respect natural phenomena and explain how this is biologically useful (NOT YET).
-Second, we break ground on the ability of GANs to learn representations of cell images and establish their performance at the task of drug compound mechanism-of-action prediction.
-We discuss that even though the representations learnt by adversarial models are currently inferior at this task, GANs provide valuable benefits over each of the competing approaches.
+This work investigates the use of GANs for the domain of cell microscopy
+imaging, particularly morphological profiling. First, we demonstrate that GANs
+are able to synthesize biologically realistic images of cells. In addition, we
+explain that GANs induce a rich semantic structure in their noise space,
+providing a powerful basis for interpretability of observed properties. Second,
+we add an encoding component to standard GANs and evaluate the quality of
+learned representations via their mechanism-of-action classification performance
+following \cite{Ljosa2013}. Even though adversarially learned representations
+are currently inferior at this task, we argue that further enhancements to GANs,
+especially biologically meaningful ones, will lead to higher quality latent
+representations.
 
-We emphasize that this work has shed light only on the tip of the iceberg of how the adversarial framework is useful for the field of computational biology.
-Due to the generality of the adversarial framework, there remain many more avenues to explore.
-For example, we hope that future work investigates recently proposed bidirectional GAN models or other, novel approaches to infering representations under the GAN umbrella.
-Finally, we note that deep neural network architectures such as those described in this work flourish with large amounts of data.
-As such, we conjecture that the performance of approaches outlined in this work will improve if and when larger biological image datasets become available.
+We emphasize that this work covers only a very small fraction of possible
+applications of GANs to the domain of microscopical images. For example, we hope
+that future work investigates BiGANs \cite{donahue2016adversarial} or other
+novel solutions to infer latent features with GANs. We believe that improved
+inference, combined with the interpretable nature of the GAN framework, may
+enable simulated experiments by performing algebra with vectors corresponding to
+cell lines, diseases or other perturbations. Finally, we note that deep learning
+generally flourishes with larger amounts of data. As such, we conjecture that
+the performance of the outlined approaches will improve with the use of larger
+datasets.
 
-\section*{References}
+\section*{Acknowledgements}
+Nick Pawlowski is supported by Microsoft Research through its PhD Scholarship
+Program and the EPSRC Centre for Doctoral Training in High Performance
+Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
 
-References follow the acknowledgments. Use unnumbered first-level
-heading for the references. Any choice of citation style is acceptable
-as long as you are consistent. It is permissible to reduce the font
-size to \verb+small+ (9 point) when listing the references. {\bf
-  Remember that you can go over 8 pages as long as the subsequent ones contain
-  \emph{only} cited references.}
-\medskip
-
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
-for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
-T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
-  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
-  Exploring Realistic Neural Models with the GEneral NEural SImulation
-  System.}  New York: TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
-learning and recall at excitatory recurrent synapses and cholinergic
-modulation in rat hippocampal region CA3. {\it Journal of
-  Neuroscience} {\bf 15}(7):5249-5262.
+\bibliographystyle{plain}
+\bibliography{lit}
 

### 2017-10-16 cytogan/train/bbbc021.py
Changed script to make gifs

@@ -301,8 +301,6 @@ with common.get_session(options.gpus) as session:
     if options.interpolate_samples is not None:
         start, end = np.random.randn(2, options.interpolate_samples[0],
                                      model.noise_size)
-        start = start[[7, 9, 13], :]
-        end = end[[7, 9, 13], :]
         if conditional_shape:
             labels = cell_data.sample_labels(options.interpolate_samples[0])
             labels = np.array(list(labels))
@@ -316,6 +314,7 @@ with common.get_session(options.gpus) as session:
             len(start),
             options.interpolate_samples[1],
             options.interpolation_method,
+            options.store_interpolation_frames,
             conditional=labels,
             gray=True,
             save_to=options.figure_dir)

### 2017-10-16 cytogan/train/common.py

@@ -60,6 +60,7 @@ def make_parser(name):
     parser.add_argument('-l', '--latent-samples', type=int)
     parser.add_argument('-g', '--generative-samples', type=int)
     parser.add_argument('-i', '--interpolate-samples', type=int, nargs=2)
+    parser.add_argument('--store-interpolation-frames', action='store_true')
     parser.add_argument(
         '--interpolation-method', default='linear', choices=('linear', 'slerp'))
     parser.add_argument('--gpus', type=int, nargs='+')

### 2017-10-16 cytogan/train/visualize.py

@@ -1,10 +1,10 @@
 import os
-import time
 
 import matplotlib.pyplot as plot
 import numpy as np
 import seaborn
 import sklearn.manifold
+import scipy.misc
 
 from cytogan.extra import logs
 
@@ -128,6 +128,7 @@ def interpolation(model,
                   number_of_interpolations,
                   interpolation_length,
                   method,
+                  store_interpolation_frames=False,
                   conditional=None,
                   gray=False,
                   save_to=None,
@@ -152,8 +153,18 @@ def interpolation(model,
     if _is_grayscale(images):
         images = _make_rgb(images)
 
-    figure = plot.figure(figsize=(8, k))
-    # figure.suptitle(title)
+    if store_interpolation_frames:
+        assert save_to is not None
+        for n, series in enumerate(np.split(images, k)):
+            folder = os.path.join(save_to, 'interpolation', str(n))
+            if not os.path.exists(folder):
+                os.makedirs(folder)
+            log.info('Storing interpolation frames to %s', folder)
+            for i, image in enumerate(series):
+                path = os.path.join(folder, '{0}.png'.format(i))
+                scipy.misc.imsave(path, image)
+
+    plot.figure(figsize=(8, k))
     for index, image in enumerate(images):
         _plot_image_tile(k, interpolation_length, index, image, gray)
 

### 2017-10-16 scripts/make_gif.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python3
+
+import argparse
+import os
+import subprocess
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-o', '--output', default='animated.gif')
+parser.add_argument('-r', '--skip-rate', type=int, default=1)
+parser.add_argument('-d', '--delay', type=int, default=10)
+parser.add_argument('files', nargs='+')
+options = parser.parse_args()
+
+options.files.sort(key=lambda p: int(os.path.basename(p).split('.')[0]))
+files = options.files[::options.skip_rate]
+files = ' '.join(files)
+
+command = 'convert -background white -alpha remove -duplicate 1,-2-1 '
+command += f'-loop 0 -delay {options.delay} '
+command += f'{files} {options.output}'
+
+print(command)
+subprocess.run(command.split())

### 2017-10-16 scripts/make_gif.sh

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-convert -background white -alpha remove -layers OptimizePlus -delay 10 \
-  "$1/%d.png[0-$(($(ls $1 | wc -l)-1))]" -loop 0 ${2:-animated.gif}

### 2017-10-16 cytogan/train/visualize.py
Bug in latent space visualizatoin

@@ -76,15 +76,15 @@ def latent_space(latent_vectors,
     for p in perplexity:
         reduction = sklearn.manifold.TSNE(
             n_components=2, perplexity=p, init='pca', verbose=1)
-        latent_vectors = reduction.fit_transform(latent_vectors)
+        transformed_vectors = reduction.fit_transform(latent_vectors)
 
         figure = plot.figure(figsize=(12, 10))
         subject_title = ' ({0})'.format(subject) if subject else ''
         subject_title += ' | P = {0}'.format(p)
         figure.suptitle('Latent Space{0}'.format(subject_title))
         plot.scatter(
-            latent_vectors[:, 0],
-            latent_vectors[:, 1],
+            transformed_vectors[:, 0],
+            transformed_vectors[:, 1],
             c=labels,
             lw=point_sizes,
             cmap=plot.cm.Spectral)

### 2017-10-17 cytogan/models/__init__.py
Making progress on OrbitalGAN

@@ -1,3 +1,13 @@
-MODELS = ['ae', 'conv_ae', 'vae', 'infogan', 'dcgan', 'lsgan', 'wgan', 'began']
+MODELS = [
+    'ae',
+    'conv_ae',
+    'vae',
+    'infogan',
+    'dcgan',
+    'lsgan',
+    'wgan',
+    'began',
+    'ogan',
+]
 for _gan_name in ('dcgan', 'lsgan', 'wgan', 'began'):
     MODELS.append('c-{0}'.format(_gan_name))

### 2017-10-17 cytogan/models/gan.py

@@ -3,13 +3,7 @@ import numpy as np
 import tensorflow as tf
 from keras.layers import Input
 
-from cytogan.models import model
-
-
-def _merge_summaries(scope):
-    scope = 'summary/{0}'.format(scope)
-    summaries = tf.get_collection(tf.GraphKeys.SUMMARIES, scope=scope)
-    return tf.summary.merge(summaries) if summaries else None
+from cytogan.models import model, util
 
 
 def get_conditional_inputs(scopes, conditional_shape):
@@ -53,8 +47,8 @@ class GAN(model.Model):
 
         super(GAN, self).__init__(learning, session)
 
-        self.generator_summary = _merge_summaries('G')
-        self.discriminator_summary = _merge_summaries('D')
+        self.generator_summary = util.merge_summaries('G')
+        self.discriminator_summary = util.merge_summaries('D')
         if (self.generator_summary is not None
                 and self.discriminator_summary is not None):
             self.summary = tf.summary.merge(

### 2017-10-17 cytogan/models/orbital_gan.py

@@ -1,0 +1,114 @@
+# 1. For each compound in batch:
+#   1. Advance moving average
+#   2. Penalize cosine distance between points of one compound and average
+#   3. Find two nearest neighbors, slightly penalize closeness
+# 2. For each concentration in batch:
+#   1. Advance moving average
+#   2. Penalize error in norm of one concentration and average
+#   3. Find two nearest neighbors, slightly penalize closeness
+
+import collections
+
+import numpy as np
+import tensorflow as tf
+import keras.backend as K
+
+from keras.layers import Input
+
+from cytogan.models import lsgan
+
+Hyper = collections.namedtuple('Hyper', [
+    'image_shape',
+    'generator_filters',
+    'discriminator_filters',
+    'generator_strides',
+    'discriminator_strides',
+    'latent_size',
+    'noise_size',
+    'initial_shape',
+    'number_of_angles',
+    'number_of_radii',
+    'origin_label',
+])
+
+
+class OrbitalGAN(lsgan.LSGAN):
+    def __init__(self, hyper, learning, session):
+        super(OrbitalGAN, self).__init__(hyper, learning, session)
+
+    def _train_discriminator(self, fake_images, real_images, labels,
+                             with_summary):
+        discriminator_labels = np.concatenate(
+            [np.zeros(len(fake_images)),
+             np.ones(len(real_images))], axis=0)
+        images = np.concatenate([fake_images, real_images], axis=0)
+        fetches = [self.optimizer['D'], self.loss['D']]
+        if with_summary and self.discriminator_summary is not None:
+            fetches.append(self.discriminator_summary)
+
+        angle_labels = labels
+        # angle_labels, radius_labels = tf.split(
+        #     labels, [self.number_of_angles, self.radius_label_size])
+
+        feed_dict = {
+            self.batch_size: [len(fake_images)],
+            self.images: images,
+            self.discriminator_labels: discriminator_labels,
+            self.angle_labels: angle_labels,
+            K.learning_phase(): 1,
+        }
+        return self.session.run(fetches, feed_dict)[1:]
+
+    def _define_graph(self):
+        super(OrbitalGAN, self)._define_graph()
+
+        # To avoid ambguity -- these are the output (e.g. probability) labels
+        self.discriminator_labels = self.labels
+
+        batch_size = tf.cast(tf.squeeze(self.batch_size), tf.int32)
+        real_latent = self.latent[batch_size:]
+
[…]

### 2017-10-17 cytogan/models/util.py

@@ -1,0 +1,6 @@
+import tensorflow as tf
+
+def merge_summaries(scope):
+    scope = 'summary/{0}'.format(scope)
+    summaries = tf.get_collection(tf.GraphKeys.SUMMARIES, scope=scope)
+    return tf.summary.merge(summaries) if summaries else None

### 2017-10-17 cytogan/train/common.py

@@ -66,6 +66,7 @@ def make_parser(name):
     parser.add_argument('--gpus', type=int, nargs='+')
     parser.add_argument('--show-figures', action='store_true')
     parser.add_argument('--skip-training', action='store_true')
+    parser.add_argument('--with-labels', action='store_true')
     parser.add_argument(
         '--summary-freq', type=Frequency, default=Frequency('20'))
     parser.add_argument(
@@ -105,6 +106,7 @@ def parse_args(parser):
     if options.model.startswith('c-'):
         options.model = options.model[2:]
         options.conditional = True
+        options.with_labels = True
     else:
         options.conditional = False
 

### 2017-10-17 cytogan/train/mnist.py

@@ -6,7 +6,7 @@ from tensorflow.examples.tutorials import mnist
 
 from cytogan.extra import distributions, logs, misc
 from cytogan.models import (ae, began, conv_ae, dcgan, infogan, lsgan, model,
-                            vae, wgan)
+                            vae, wgan, orbital_gan)
 from cytogan.train import common, trainer, visualize
 
 parser = common.make_parser('cytogan-mnist')
@@ -30,11 +30,12 @@ else:
 
 def get_batch(n):
     images, labels = data.train.next_batch(n)
-    images = images.reshape((-1, ) + image_shape)
+    batch = [images.reshape((-1, ) + image_shape)]
+    if options.with_labels:
+        batch.append(labels.argmax(axis=1))
     if options.conditional:
-        return (images, labels.reshape(-1, *conditional_shape))
-    else:
-        return images
+        batch.append(labels)
+    return batch[0] if len(batch) == 1 else batch
 
 
 learning = model.Learning(options.lr, options.lr_decay, options.lr_decay_steps
@@ -104,6 +105,20 @@ elif options.model == 'infogan':
         probability_loss='bce',
         continuous_loss='bce')
     Model = infogan.InfoGAN
+elif options.model == 'ogan':
+    hyper = orbital_gan.Hyper(
+        image_shape,
+        generator_filters=(128, 64, 32, 16),
+        generator_strides=(1, 2, 2, 1),
+        discriminator_filters=(128, 64, 32, 16),
+        discriminator_strides=(1, 2, 2, 2),
+        latent_size=100,
+        noise_size=100,
+        initial_shape=(7, 7),
+        number_of_angles=10,
+        number_of_radii=None,
+        origin_label=0)
+    Model = orbital_gan.OrbitalGAN
 
 log.debug('Hyperparameters:\n%s', misc.namedtuple_to_string(hyper))
 
@@ -151,7 +166,7 @@ with common.get_session(options.gpus) as session:
         original_images, labels = data.test.next_batch(options.latent_samples)
         original_images = original_images.reshape(-1, 28, 28, 1)
         latent_vectors = model.encode(original_images)
-        labels = np.argmax(labels, axis=1)
+        labels = np.equal(np.argmax(labels, axis=1), 0).astype(np.int32)
         visualize.latent_space(
             latent_vectors, labels, save_to=options.figure_dir)
 

### 2017-10-17 cytogan/models/orbital_gan.py
Debugging vector norm

@@ -87,7 +87,7 @@ class OrbitalGAN(lsgan.LSGAN):
         with tf.control_dependencies([update_ema_op]):
             origin_mask = tf.equal(self.angle_labels, self.origin_label)
             origin_vectors = tf.boolean_mask(real_latent, origin_mask)
-            self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=0))
+            self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
         self.loss['D'] += self.origin_norm
         # self.loss['O'] = origin_norm

### 2017-10-17 cytogan/train/visualize.py

@@ -68,6 +68,9 @@ def latent_space(latent_vectors,
     assert np.ndim(latent_vectors) == 2
     log.info('Plotting latent space for %d vectors', len(latent_vectors))
 
+    for v, l in zip(latent_vectors, labels):
+        print('{0} {1}'.format(np.linalg.norm(v), l))
+
     if perplexity is None:
         perplexity = [2, 3, 4, 5] + list(range(10, 21)) + [30, 50, 70, 90]
     if isinstance(perplexity, int):

### 2017-10-17 cytogan/models/orbital_gan.py
Testing origin centering

@@ -79,8 +79,8 @@ class OrbitalGAN(lsgan.LSGAN):
             self.angle_means.append(mean)
             self.angle_variances.append(variance)
 
-        ema = tf.train.ExponentialMovingAverage(decay=0.99)
-        update_ema_op = ema.apply(self.angle_means + self.angle_variances)
+        self.ema = tf.train.ExponentialMovingAverage(decay=0.9999)
+        update_ema_op = self.ema.apply(self.angle_means + self.angle_variances)
 
         # self.radius_labels = Input(shape=[self.radius_label_shape])
 
@@ -89,7 +89,7 @@ class OrbitalGAN(lsgan.LSGAN):
             origin_vectors = tf.boolean_mask(real_latent, origin_mask)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
-        self.loss['D'] += self.origin_norm
+        self.loss['D'] += 0.1 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):
@@ -111,4 +111,5 @@ class OrbitalGAN(lsgan.LSGAN):
         with K.name_scope('summary/D'):
             tf.summary.scalar('origin_norm', self.origin_norm)
             for index, variance in enumerate(self.angle_variances):
-                tf.summary.scalar('angle-variance-{0}'.format(index), tf.norm(variance))
+                tf.summary.scalar('angle-variance-{0}'.format(index),
+                                  tf.norm(self.ema.average(variance)))

### 2017-10-17 cytogan/train/visualize.py

@@ -33,7 +33,7 @@ def _save_figure(folder, filename):
     if not os.path.exists(folder):
         os.makedirs(folder)
     log.info('Saving %s', path)
-    plot.savefig(path, bbox_inches='tight', pad_inches=0)
+    plot.savefig(path)
 
 
 def reconstructions(model,
@@ -172,7 +172,6 @@ def interpolation(model,
         _plot_image_tile(k, interpolation_length, index, image, gray)
 
     plot.subplots_adjust(left=0.1, right=0.9, top=0.6, bottom=0.2)
-    plot.tight_layout()
 
     if save_to is not None:
         filename = '{0}-interpolation.png'.format(method)

### 2017-10-17 cytogan/models/orbital_gan.py
Adding more regularization

@@ -89,7 +89,7 @@ class OrbitalGAN(lsgan.LSGAN):
             origin_vectors = tf.boolean_mask(real_latent, origin_mask)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
-        self.loss['D'] += 0.1 * self.origin_norm
+        self.loss['D'] += 1 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):

### 2017-10-17 cytogan/models/orbital_gan.py
Trying my own boolean mask

@@ -86,7 +86,7 @@ class OrbitalGAN(lsgan.LSGAN):
 
         with tf.control_dependencies([update_ema_op]):
             origin_mask = tf.equal(self.angle_labels, self.origin_label)
-            origin_vectors = tf.boolean_mask(real_latent, origin_mask)
+            origin_vectors = real_latent * tf.cast(origin_mask, tf.float32)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
         self.loss['D'] += 1 * self.origin_norm

### 2017-10-17 cytogan/models/orbital_gan.py
p[weuV       DJNSCKLX/AUI;CFDUHAUDPHFGIUDHFHAPDH

@@ -85,11 +85,15 @@ class OrbitalGAN(lsgan.LSGAN):
         # self.radius_labels = Input(shape=[self.radius_label_shape])
 
         with tf.control_dependencies([update_ema_op]):
-            origin_mask = tf.equal(self.angle_labels, self.origin_label)
-            origin_vectors = real_latent * tf.cast(origin_mask, tf.float32)
-            self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
+            origin_mask = tf.cast(
+                tf.equal(self.angle_labels, self.origin_label), tf.float32)
+            origin_vectors = real_latent * tf.expand_dims(origin_mask, 1)
+            print(origin_vectors)
+            assert len(origin_vectors.shape.as_list()) == 2
+            self.origin_norm = tf.reduce_mean(
+                tf.norm(origin_vectors, axis=1) + 1e-8)
 
-        self.loss['D'] += 1 * self.origin_norm
+        self.loss['D'] += 0.1 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):

### 2017-10-17 cytogan/models/orbital_gan.py
Revert "p[weuV       DJNSCKLX/AUI;CFDUHAUDPHFGIUDHFHAPDH"
This reverts commit 0a0287daef572b4b524a676cfc263cf4dd4522eb.

@@ -85,15 +85,11 @@ class OrbitalGAN(lsgan.LSGAN):
         # self.radius_labels = Input(shape=[self.radius_label_shape])
 
         with tf.control_dependencies([update_ema_op]):
-            origin_mask = tf.cast(
-                tf.equal(self.angle_labels, self.origin_label), tf.float32)
-            origin_vectors = real_latent * tf.expand_dims(origin_mask, 1)
-            print(origin_vectors)
-            assert len(origin_vectors.shape.as_list()) == 2
-            self.origin_norm = tf.reduce_mean(
-                tf.norm(origin_vectors, axis=1) + 1e-8)
+            origin_mask = tf.equal(self.angle_labels, self.origin_label)
+            origin_vectors = real_latent * tf.cast(origin_mask, tf.float32)
+            self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
-        self.loss['D'] += 0.1 * self.origin_norm
+        self.loss['D'] += 1 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):

### 2017-10-17 cytogan/models/orbital_gan.py
Revert "Trying my own boolean mask"
This reverts commit 3a6802e38b062b4d0db2ca05f1f9f407c56f0b5c.

@@ -86,7 +86,7 @@ class OrbitalGAN(lsgan.LSGAN):
 
         with tf.control_dependencies([update_ema_op]):
             origin_mask = tf.equal(self.angle_labels, self.origin_label)
-            origin_vectors = real_latent * tf.cast(origin_mask, tf.float32)
+            origin_vectors = tf.boolean_mask(real_latent, origin_mask)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
         self.loss['D'] += 1 * self.origin_norm

### 2017-10-17 cytogan/models/orbital_gan.py
Revert "Adding more regularization"
This reverts commit 330b01cb9400830aa21619c11d177137e67363c4.

@@ -89,7 +89,7 @@ class OrbitalGAN(lsgan.LSGAN):
             origin_vectors = tf.boolean_mask(real_latent, origin_mask)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
-        self.loss['D'] += 1 * self.origin_norm
+        self.loss['D'] += 0.1 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):

### 2017-10-18 cytogan/metrics/profiling.py
Trying different profiling distance

@@ -14,7 +14,7 @@ def get_nearest_neighbors(examples, neighbors):
     example_matrix = np.array(list(examples))
     neighbor_matrix = np.array(list(neighbors))
     # Gives us |examples| x |neighbors| matrix.
-    distances = sklearn.metrics.pairwise.cosine_distances(
+    distances = sklearn.metrics.pairwise.euclidean_distances(
         example_matrix, neighbor_matrix)
     # Get the indices of the nearest neighbor for each test sample.
     return np.argmin(distances, axis=1)

### 2017-10-18 cytogan/models/dcgan.py

@@ -72,13 +72,7 @@ class DCGAN(gan.GAN):
 
     def _define_graph(self):
         if self.is_conditional:
-            self.conditional = gan.get_conditional_inputs(
-                ('G', 'D'), self.conditional_shape)
-            if self.conditional_embedding is not None:
-                self.conditional_embedding_layer = Dense(
-                    self.conditional_embedding,
-                    activation='relu',
-                    name='embedding')
+            self._define_conditional_units(self.conditional_shape)
 
         with K.name_scope('G'):
             self.batch_size = Input(batch_shape=[1], name='batch_size')
@@ -150,6 +144,15 @@ class DCGAN(gan.GAN):
         D = Flatten()(D)
 
         return D
+
+    def _define_conditional_units(self, conditional_shape):
+        self.conditional = gan.get_conditional_inputs(
+            ('G', 'D'), conditional_shape)
+        if self.conditional_embedding is not None:
+            self.conditional_embedding_layer = Dense(
+                self.conditional_embedding,
+                activation='relu',
+                name='embedding')
 
     def _define_generator_loss(self, probability):
         with K.name_scope('G_loss'):

### 2017-10-18 cytogan/models/orbital_gan.py

@@ -66,7 +66,11 @@ class OrbitalGAN(lsgan.LSGAN):
         self.discriminator_labels = self.labels
 
         batch_size = tf.cast(tf.squeeze(self.batch_size), tf.int32)
+
+        # with tf.control_dependencies([tf.check_numerics(self.latent, 'latent')]):
         real_latent = self.latent[batch_size:]
+
+        real_latent = tf.Print(real_latent, [real_latent])
 
         self.angle_labels = Input(batch_shape=[None], dtype=tf.int32)
 
@@ -85,11 +89,15 @@ class OrbitalGAN(lsgan.LSGAN):
         # self.radius_labels = Input(shape=[self.radius_label_shape])
 
         with tf.control_dependencies([update_ema_op]):
-            origin_mask = tf.equal(self.angle_labels, self.origin_label)
-            origin_vectors = tf.boolean_mask(real_latent, origin_mask)
+            # origin_mask = tf.equal(self.angle_labels, self.origin_label)
+            # origin_vectors = tf.boolean_mask(real_latent, origin_mask)
+            origin_mask = tf.cast(
+                tf.equal(self.angle_labels, self.origin_label), tf.float32)
+            origin_mask = tf.Print(origin_mask, [real_latent, origin_mask])
+            origin_vectors = real_latent * tf.expand_dims(origin_mask, 1)
             self.origin_norm = tf.reduce_mean(tf.norm(origin_vectors, axis=1))
 
-        self.loss['D'] += 0.1 * self.origin_norm
+        self.loss['D'] += 0.000001 * self.origin_norm
         # self.loss['O'] = origin_norm
 
     def train_on_batch(self, batch, with_summary=False):

### 2017-10-18 cytogan/metrics/profiling.py
Added more logging to profiling

@@ -7,19 +7,6 @@ from cytogan.extra import logs
 
 log = logs.get_logger(__name__)
 
-
-def get_nearest_neighbors(examples, neighbors):
-    assert len(examples) > 0
-    assert len(neighbors) > 0
-    example_matrix = np.array(list(examples))
-    neighbor_matrix = np.array(list(neighbors))
-    # Gives us |examples| x |neighbors| matrix.
-    distances = sklearn.metrics.pairwise.euclidean_distances(
-        example_matrix, neighbor_matrix)
-    # Get the indices of the nearest neighbor for each test sample.
-    return np.argmin(distances, axis=1)
-
-
 def reduce_profiles_across_treatments(dataset):
     keys = ('compound', 'concentration', 'moa')
     reduced_profiles = []
@@ -56,13 +43,18 @@ def whiten(dataset):
     dataset['profile'] = list(whitened_profiles)
 
 
-# Performs leave-one-compound-out cross-validation
-# Result is a dataframe with columns:
-# - image key
-# - compound
-# - concentration
-# - profile vector
-# - MOA
+def get_nearest_neighbors(examples, neighbors):
+    assert len(examples) > 0
+    assert len(neighbors) > 0
+    example_matrix = np.array(list(examples))
+    neighbor_matrix = np.array(list(neighbors))
+    # Gives us |examples| x |neighbors| matrix.
+    distances = sklearn.metrics.pairwise.cosine_distances(
+        example_matrix, neighbor_matrix)
+    # Get the indices of the nearest neighbor for each test sample.
+    return np.argmin(distances, axis=1)
+
+
 def score_profiles(dataset):
     accuracies = []
     labels = dataset['moa'].unique()
@@ -79,9 +71,9 @@ def score_profiles(dataset):
         test_data = dataset[test_mask]
         # All other (compound, concentration) pairs.
         training_data = dataset[~test_mask]
-        if training_data.empty:
-            continue
 
+        log.info('Finding NN of %d treatments among %d other treatments',
+                 len(test_data), len(training_data))
         neighbor_indices = get_nearest_neighbors(test_data['profile'],
                                                  training_data['profile'])
         # Get the MOAs of those nearest neighbors as our predictions.

### 2017-10-18 cytogan/models/orbital_gan.py

@@ -17,6 +17,8 @@ from keras.layers import Input
 
 from cytogan.models import lsgan
 
+tf.logging.set_verbosity(tf.logging.INFO)
+
 Hyper = collections.namedtuple('Hyper', [
     'image_shape',
     'generator_filters',
@@ -31,6 +33,15 @@ Hyper = collections.namedtuple('Hyper', [
     'origin_label',
 ])
 
+def tf_print(op, tensors):
+    def print_message(x):
+        print(x)
+        return x
+
+    prints = [tf.py_func(print_message, [tensor], tensor.dtype) for tensor in tensors]
+    with tf.control_dependencies(prints):
+        op = tf.identity(op)
+    return op
 
 class OrbitalGAN(lsgan.LSGAN):
     def __init__(self, hyper, learning, session):
@@ -67,37 +78,38 @@ class OrbitalGAN(lsgan.LSGAN):
 
         batch_size = tf.cast(tf.squeeze(self.batch_size), tf.int32)
 
-        # with tf.control_dependencies([tf.check_numerics(self.latent, 'latent')]):
         real_latent = self.latent[batch_size:]
-
-        real_latent = tf.Print(real_latent, [real_latent])
-
         self.angle_labels = Input(batch_shape=[None], dtype=tf.int32)
 
         self.angle_means, self.angle_variances = [], []
-        for partition in tf.dynamic_partition(
+        angle_partitions = tf.dynamic_partition(
                 real_latent,
                 self.angle_labels,
-                num_partitions=self.number_of_angles):
-            mean, variance = tf.nn.moments(partition, axes=[0])
+                num_partitions=self.number_of_angles)
+        for group in angle_partitions:
+            mean, variance = tf.nn.moments(group, axes=[0])
             self.angle_means.append(mean)
             self.angle_variances.append(variance)
 
         self.ema = tf.train.ExponentialMovingAverage(decay=0.9999)
         update_ema_op = self.ema.apply(self.angle_means + self.angle_variances)
 
+        for mean, vectors in zip(self.angle_means, angle_partitions):
+            moving_mean = self.ema.average(mean)
+
+
+        mean_variance = tf.reduce_mean([self.ema.average(v) for v in self.angle_variances])
+
+        self.loss['D'] += 1 * mean_variance
+
         # self.radius_labels = Input(shape=[self.radius_label_shape])
 
         with tf.control_dependencies([update_ema_op]):
[…]

### 2017-10-18 cytogan/train/visualize.py

@@ -68,9 +68,6 @@ def latent_space(latent_vectors,
     assert np.ndim(latent_vectors) == 2
     log.info('Plotting latent space for %d vectors', len(latent_vectors))
 
-    for v, l in zip(latent_vectors, labels):
-        print('{0} {1}'.format(np.linalg.norm(v), l))
-
     if perplexity is None:
         perplexity = [2, 3, 4, 5] + list(range(10, 21)) + [30, 50, 70, 90]
     if isinstance(perplexity, int):

### 2017-10-18 cytogan/train/bbbc021.py
Added ability to load profiles from disk

@@ -1,9 +1,11 @@
 #!/usr/bin/env python3
 
 import os
+
 import numpy as np
-import tensorflow as tf
+import pandas as pd
 import scipy.misc
+import tensorflow as tf
 from tqdm import tqdm
 
 from cytogan.data.cell_data import CellData
@@ -27,6 +29,7 @@ parser.add_argument('--normalize-luminance', action='store_true')
 parser.add_argument('--whiten-profiles', action='store_true')
 parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
+parser.add_argument('--load-profiles')
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
@@ -184,23 +187,34 @@ with common.get_session(options.gpus) as session:
 
     if not options.skip_evaluation:
         log.info('Starting Evaluation')
-        keys, profiles = [], []
-        batch_generator = cell_data.batches_of_size(options.batch_size)
-        try:
-            for batch_keys, images in tqdm(batch_generator, unit=' batches'):
-                profiles.append(model.encode(images))
-                keys += batch_keys
-        except KeyboardInterrupt:
-            pass
 
-        profiles = np.concatenate(profiles, axis=0)
-        log.info('Generated %d profiles', len(profiles))
+        if options.load_profiles is None:
+            keys, profiles = [], []
+            batch_generator = cell_data.batches_of_size(options.batch_size)
+            try:
+                for batch_keys, images in tqdm(
+                        batch_generator, unit=' batches'):
+                    profiles.append(model.encode(images))
+                    keys += batch_keys
+            except KeyboardInterrupt:
+                pass
 
-        dataset = cell_data.create_dataset_from_profiles(keys, profiles)
-        log.info('Matching {0:,} profiles to {1} MOAs'.format(
-            len(dataset), len(dataset.moa.unique())))
+            profiles = np.concatenate(profiles, axis=0)
+            log.info('Generated %d profiles', len(profiles))
 
-        if options.save_profiles:
+            dataset = cell_data.create_dataset_from_profiles(keys, profiles)
+            log.info('Matching {0:,} profiles to {1} MOAs'.format(
+                len(dataset), len(dataset.moa.unique())))
+        else:
+            log.info('Loading profiles from %s', options.load_profiles)
+            dataset = pd.read_csv(
+                options.load_profiles,
[…]

### 2017-10-18 cytogan/train/bbbc021.py
Fixed profile loading

@@ -42,6 +42,7 @@ if options.save_profiles:
         os.makedirs(options.profiles_dir)
 
     def save_profiles(profiles, filename):
+        filename += '.gz'
         log.info('Storing %s to disk', filename)
         path = os.path.join(options.profiles_dir, filename)
         profiles.to_csv(
@@ -209,7 +210,6 @@ with common.get_session(options.gpus) as session:
             log.info('Loading profiles from %s', options.load_profiles)
             dataset = pd.read_csv(
                 options.load_profiles,
-                header=True,
                 compression='gzip',
                 encoding='ascii',
                 chunksize=100000)

### 2017-10-18 cytogan/train/bbbc021.py
Removed chunksize from profile loading

@@ -209,10 +209,8 @@ with common.get_session(options.gpus) as session:
         else:
             log.info('Loading profiles from %s', options.load_profiles)
             dataset = pd.read_csv(
-                options.load_profiles,
-                compression='gzip',
-                encoding='ascii',
-                chunksize=100000)
+                options.load_profiles, compression='gzip', encoding='ascii')
+            log.info('Found %s profiles', len(dataset))
 
         if options.save_profiles and not options.load_profiles:
             log.info('Storing profiles to disk')

### 2017-10-18 cytogan/train/bbbc021.py
Added ability to load mean profiles

@@ -30,6 +30,7 @@ parser.add_argument('--whiten-profiles', action='store_true')
 parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
 parser.add_argument('--load-profiles')
+parser.add_argument('--load-average-profiles')
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
@@ -189,7 +190,9 @@ with common.get_session(options.gpus) as session:
     if not options.skip_evaluation:
         log.info('Starting Evaluation')
 
-        if options.load_profiles is None:
+        loaded_profiles = (options.load_profiles
+                           or options.load_collapsed_profiles)
+        if not loaded_profiles:
             keys, profiles = [], []
             batch_generator = cell_data.batches_of_size(options.batch_size)
             try:
@@ -206,15 +209,16 @@ with common.get_session(options.gpus) as session:
             dataset = cell_data.create_dataset_from_profiles(keys, profiles)
             log.info('Matching {0:,} profiles to {1} MOAs'.format(
                 len(dataset), len(dataset.moa.unique())))
-        else:
+
+            if options.save_profiles:
+                log.info('Storing profiles to disk')
+                save_profiles(dataset, 'profiles.csv')
+
+        if options.load_profiles:
             log.info('Loading profiles from %s', options.load_profiles)
             dataset = pd.read_csv(
                 options.load_profiles, compression='gzip', encoding='ascii')
             log.info('Found %s profiles', len(dataset))
-
-        if options.save_profiles and not options.load_profiles:
-            log.info('Storing profiles to disk')
-            save_profiles(dataset, 'profiles.csv')
 
         if options.whiten_profiles:
             profiling.whiten(dataset)
@@ -222,15 +226,25 @@ with common.get_session(options.gpus) as session:
             if options.save_profiles:
                 save_profiles(dataset, 'whitened.csv')
 
-        # The DMSO (control) should not participate in the MOA classification.
-        dataset = dataset[dataset['compound'] != 'DMSO']
-        treatment_profiles = profiling.reduce_profiles_across_treatments(
-            dataset)
-        log.info('Reduced dataset from %d to %d profiles for each treatment',
-                 len(dataset), len(treatment_profiles))
+        if options.load_collapsed_profiles:
+            log.info('Loading collapsed profiles')
+            treatment_profiles = pd.read_csv(
[…]

### 2017-10-18 cytogan/train/bbbc021.py
s/averaged/collapsed

@@ -30,7 +30,7 @@ parser.add_argument('--whiten-profiles', action='store_true')
 parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
 parser.add_argument('--load-profiles')
-parser.add_argument('--load-average-profiles')
+parser.add_argument('--load-collapsed-profiles')
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
@@ -43,7 +43,7 @@ if options.save_profiles:
         os.makedirs(options.profiles_dir)
 
     def save_profiles(profiles, filename):
-        filename += '.gz'
+        filename += '.csv.gz'
         log.info('Storing %s to disk', filename)
         path = os.path.join(options.profiles_dir, filename)
         profiles.to_csv(
@@ -212,7 +212,7 @@ with common.get_session(options.gpus) as session:
 
             if options.save_profiles:
                 log.info('Storing profiles to disk')
-                save_profiles(dataset, 'profiles.csv')
+                save_profiles(dataset, 'profiles')
 
         if options.load_profiles:
             log.info('Loading profiles from %s', options.load_profiles)
@@ -224,7 +224,7 @@ with common.get_session(options.gpus) as session:
             profiling.whiten(dataset)
             log.info('Whitened data')
             if options.save_profiles:
-                save_profiles(dataset, 'whitened.csv')
+                save_profiles(dataset, 'whitened')
 
         if options.load_collapsed_profiles:
             log.info('Loading collapsed profiles')
@@ -244,7 +244,7 @@ with common.get_session(options.gpus) as session:
                 len(dataset), len(treatment_profiles))
 
             if options.save_profiles:
-                save_profiles(treatment_profiles, 'treatments.csv')
+                save_profiles(treatment_profiles, 'treatments')
 
         confusion_matrix, accuracy = profiling.score_profiles(
             treatment_profiles)

### 2017-10-18 cytogan/metrics/profiling.py
Better parsing of loaded profiles

@@ -1,3 +1,5 @@
+import re
+
 import numpy as np
 import pandas as pd
 import scipy.linalg
@@ -6,6 +8,28 @@ import sklearn.metrics.pairwise
 from cytogan.extra import logs
 
 log = logs.get_logger(__name__)
+
+
+def save_profiles(path, profiles):
+    profiles.to_csv(
+        path,
+        header=True,
+        compression='gzip',
+        encoding='ascii',
+        chunksize=100000)
+
+
+def load_profiles(path):
+    log.info('Loading profiles from %s', path)
+    data = pd.read_csv(path, compression='gzip', encoding='ascii')
+    parsed_profiles = []
+    for p in data['profile']:
+        values = re.sub(r'[\[\]\n]', '', p).split()
+        parsed_profiles.append(np.array([float(x) for x in values]))
+    data.profile = parsed_profiles
+
+    return data
+
 
 def reduce_profiles_across_treatments(dataset):
     keys = ('compound', 'concentration', 'moa')

### 2017-10-18 cytogan/train/bbbc021.py

@@ -42,16 +42,11 @@ if options.save_profiles:
     if not os.path.exists(options.profiles_dir):
         os.makedirs(options.profiles_dir)
 
-    def save_profiles(profiles, filename):
-        filename += '.csv.gz'
+    def save_profiles(profiles, prefix):
+        filename = '{0}.csv.gz'.format(prefix)
         log.info('Storing %s to disk', filename)
         path = os.path.join(options.profiles_dir, filename)
-        profiles.to_csv(
-            path,
-            header=True,
-            compression='gzip',
-            encoding='ascii',
-            chunksize=100000)
+        profiling.save_profiles(path, profiles)
 
 
 log = logs.get_root_logger(options.log_file)
@@ -215,9 +210,7 @@ with common.get_session(options.gpus) as session:
                 save_profiles(dataset, 'profiles')
 
         if options.load_profiles:
-            log.info('Loading profiles from %s', options.load_profiles)
-            dataset = pd.read_csv(
-                options.load_profiles, compression='gzip', encoding='ascii')
+            dataset = profiling.load_profiles(options.load_profiles)
             log.info('Found %s profiles', len(dataset))
 
         if options.whiten_profiles:
@@ -227,11 +220,8 @@ with common.get_session(options.gpus) as session:
                 save_profiles(dataset, 'whitened')
 
         if options.load_collapsed_profiles:
-            log.info('Loading collapsed profiles')
-            treatment_profiles = pd.read_csv(
-                options.load_collapsed_profiles,
-                compression='gzip',
-                encoding='ascii')
+            treatment_profiles = profiling.load_profiles(
+                options.load_collapsed_profiles)
             log.info('Found %d collapsed profiles', len(treatment_profiles))
         else:
             log.info('Collapsing profiles across treatments')

### 2017-10-18 cytogan/metrics/profiling.py
Debugging profiling

@@ -96,8 +96,12 @@ def score_profiles(dataset):
         # All other (compound, concentration) pairs.
         training_data = dataset[~test_mask]
 
-        log.info('Finding NN of %d treatments among %d other treatments',
-                 len(test_data), len(training_data))
+        print(test_data)
+        log.info(
+            'Finding NN for concentrations (%s) among %d other treatments',
+            ','.join(test_data['concentration']), len(training_data))
+        print(training_data)
+        print('-' * 100)
         neighbor_indices = get_nearest_neighbors(test_data['profile'],
                                                  training_data['profile'])
         # Get the MOAs of those nearest neighbors as our predictions.

### 2017-10-18 cytogan/metrics/profiling.py
Forgot to map(str)

@@ -99,7 +99,7 @@ def score_profiles(dataset):
         print(test_data)
         log.info(
             'Finding NN for concentrations (%s) among %d other treatments',
-            ','.join(test_data['concentration']), len(training_data))
+            ','.join(map(str, test_data['concentration'])), len(training_data))
         print(training_data)
         print('-' * 100)
         neighbor_indices = get_nearest_neighbors(test_data['profile'],

### 2017-10-18 cytogan/metrics/profiling.py
Showing top k MOAs

@@ -75,8 +75,9 @@ def get_nearest_neighbors(examples, neighbors):
     # Gives us |examples| x |neighbors| matrix.
     distances = sklearn.metrics.pairwise.cosine_distances(
         example_matrix, neighbor_matrix)
+
     # Get the indices of the nearest neighbor for each test sample.
-    return np.argmin(distances, axis=1)
+    return distances, np.argmin(distances, axis=1)
 
 
 def score_profiles(dataset):
@@ -96,17 +97,25 @@ def score_profiles(dataset):
         # All other (compound, concentration) pairs.
         training_data = dataset[~test_mask]
 
-        print(test_data)
+        concentrations_string = ', '.join(map(str, test_data['concentration']))
         log.info(
-            'Finding NN for concentrations (%s) among %d other treatments',
-            ','.join(map(str, test_data['concentration'])), len(training_data))
-        print(training_data)
-        print('-' * 100)
-        neighbor_indices = get_nearest_neighbors(test_data['profile'],
-                                                 training_data['profile'])
+            'Finding NN for %d concentrations (%s) among %d other treatments',
+            len(test_data), concentrations_string, len(training_data))
+
+        distances, nearest = get_nearest_neighbors(test_data['profile'],
+                                                   training_data['profile'])
+
+        top_k = np.argsort(distances, axis=1)[:, :5]
+        for x, y in zip(test_data, top_k):
+            top_k_moas = training_data['moa'].iloc[y]
+            print('{0}/{1} ({2}): {3}'.format(
+                x['compound'],
+                x['concentration'],
+                x['moa'],
+                tuple(top_k_moas), ))
+
         # Get the MOAs of those nearest neighbors as our predictions.
-        predicted_labels = np.array(
-            training_data['moa'].iloc[neighbor_indices])
+        predicted_labels = np.array(training_data['moa'].iloc[nearest])
         actual_labels = np.array(test_data['moa'])
         assert actual_labels.shape == predicted_labels.shape
         accuracy = np.mean(predicted_labels == actual_labels)

### 2017-10-18 cytogan/metrics/profiling.py
Fixing profiling

@@ -106,13 +106,10 @@ def score_profiles(dataset):
                                                    training_data['profile'])
 
         top_k = np.argsort(distances, axis=1)[:, :5]
-        for x, y in zip(test_data, top_k):
+        for c, m, y in zip(test_data.concentration, test_data.moa, top_k):
             top_k_moas = training_data['moa'].iloc[y]
-            print('{0}/{1} ({2}): {3}'.format(
-                x['compound'],
-                x['concentration'],
-                x['moa'],
-                tuple(top_k_moas), ))
+            print('{0}/{1} ({2}): {3}'.format(holdout_compound, c, m,
+                                              tuple(top_k_moas)))
 
         # Get the MOAs of those nearest neighbors as our predictions.
         predicted_labels = np.array(training_data['moa'].iloc[nearest])

### 2017-10-18 cytogan/metrics/profiling.py
Figuring out confusion matrix

@@ -108,8 +108,11 @@ def score_profiles(dataset):
         top_k = np.argsort(distances, axis=1)[:, :5]
         for c, m, y in zip(test_data.concentration, test_data.moa, top_k):
             top_k_moas = training_data['moa'].iloc[y]
+            top_k_strings = []
+            for i, j in zip(top_k_moas, y):
+                top_k_strings.append('{0} ({1:.4f})'.format(i, distances[j]))
             print('{0}/{1} ({2}): {3}'.format(holdout_compound, c, m,
-                                              tuple(top_k_moas)))
+                                              ', '.join(top_k_strings)))
 
         # Get the MOAs of those nearest neighbors as our predictions.
         predicted_labels = np.array(training_data['moa'].iloc[nearest])
@@ -119,6 +122,8 @@ def score_profiles(dataset):
         log.info('Accuracy for %s is %.3f', holdout_compound, accuracy)
         accuracies.append(accuracy)
 
-        confusion_matrix.loc[actual_labels, predicted_labels] += 1
+        for actual in actual_labels:
+            for predicted in predicted_labels:
+                confusion_matrix[actual][predicted] += 1
 
     return confusion_matrix, np.mean(accuracies)

### 2017-10-18 cytogan/train/bbbc021.py

@@ -31,6 +31,7 @@ parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
 parser.add_argument('--load-profiles')
 parser.add_argument('--load-collapsed-profiles')
+parser.add_argument('--tsne-perplexity', type=int)
 parser.add_argument('--vector-distance', action='store_true')
 parser.add_argument('--concentration-only-labels', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
@@ -267,6 +268,7 @@ with common.get_session(options.gpus) as session:
                 np.array(list(latent_vectors)),
                 indices,
                 point_sizes=np.array(list(point_sizes)),
+                perplexity=options.tsne_perplexity,
                 save_to=options.figure_dir,
                 subject='Concentrations')
 
@@ -278,6 +280,7 @@ with common.get_session(options.gpus) as session:
                 np.array(list(latent_vectors)),
                 indices,
                 point_sizes=np.array(list(point_sizes)),
+                perplexity=options.tsne_perplexity,
                 save_to=options.figure_dir,
                 subject='MOA')
 
@@ -306,6 +309,7 @@ with common.get_session(options.gpus) as session:
         visualize.latent_space(
             latent_vectors,
             indices,
+            perplexity=options.tsne_perplexity,
             save_to=options.figure_dir,
             subject='Cells')
 

### 2017-10-18 cytogan/metrics/profiling.py
Fixing profiling

@@ -106,13 +106,16 @@ def score_profiles(dataset):
                                                    training_data['profile'])
 
         top_k = np.argsort(distances, axis=1)[:, :5]
+        index = 0
         for c, m, y in zip(test_data.concentration, test_data.moa, top_k):
             top_k_moas = training_data['moa'].iloc[y]
             top_k_strings = []
             for i, j in zip(top_k_moas, y):
-                top_k_strings.append('{0} ({1:.4f})'.format(i, distances[j]))
+                distance = distances[index, j]
+                top_k_strings.append('{0} ({1:.4f})'.format(i, distance))
             print('{0}/{1} ({2}): {3}'.format(holdout_compound, c, m,
                                               ', '.join(top_k_strings)))
+            index += 1
 
         # Get the MOAs of those nearest neighbors as our predictions.
         predicted_labels = np.array(training_data['moa'].iloc[nearest])

### 2017-10-18 cytogan/metrics/profiling.py
Added colorbar to MOA latent space

@@ -67,6 +67,20 @@ def whiten(dataset):
     dataset['profile'] = list(whitened_profiles)
 
 
+def log_top_k(compound, distances, test_data, training_data, k=3):
+    top_k = np.argsort(distances, axis=1)[:, :k]
+    concentrations = test_data.concentration
+    moas = test_data.moa
+    for i, (c, m, y) in enumerate(zip(concentrations, moas, top_k)):
+        top_k_moas = training_data['moa'].iloc[y]
+        top_k_strings = []
+        for s, j in zip(top_k_moas, y):
+            top_k_strings.append('{0} ({1:.8f})'.format(s, distances[i, j]))
+        top_k_string = ', '.join(top_k_strings)
+        log.info('Top %d neighbors for %s/%.4f (%s): %s'.format(
+            k, compound, c, m, top_k_string))
+
+
 def get_nearest_neighbors(examples, neighbors):
     assert len(examples) > 0
     assert len(neighbors) > 0
@@ -105,17 +119,7 @@ def score_profiles(dataset):
         distances, nearest = get_nearest_neighbors(test_data['profile'],
                                                    training_data['profile'])
 
-        top_k = np.argsort(distances, axis=1)[:, :5]
-        index = 0
-        for c, m, y in zip(test_data.concentration, test_data.moa, top_k):
-            top_k_moas = training_data['moa'].iloc[y]
-            top_k_strings = []
-            for i, j in zip(top_k_moas, y):
-                distance = distances[index, j]
-                top_k_strings.append('{0} ({1:.4f})'.format(i, distance))
-            print('{0}/{1} ({2}): {3}'.format(holdout_compound, c, m,
-                                              ', '.join(top_k_strings)))
-            index += 1
+        log_top_k(holdout_compound, distances, test_data, training_data)
 
         # Get the MOAs of those nearest neighbors as our predictions.
         predicted_labels = np.array(training_data['moa'].iloc[nearest])

### 2017-10-18 cytogan/train/bbbc021.py

@@ -256,6 +256,7 @@ with common.get_session(options.gpus) as session:
                 np.array(list(latent_vectors)),
                 indices,
                 point_sizes=np.array(list(point_sizes)),
+                perplexity=options.tsne_perplexity,
                 save_to=options.figure_dir,
                 subject='Compounds')
 
@@ -273,7 +274,7 @@ with common.get_session(options.gpus) as session:
                 subject='Concentrations')
 
         if options.latent_moa:
-            _, indices = cell_data.get_moa_indices(treatment_profiles)
+            moa_names, indices = cell_data.get_moa_indices(treatment_profiles)
             latent_vectors = treatment_profiles['profile']
             point_sizes = treatment_profiles.groupby('compound').cumcount()
             visualize.latent_space(
@@ -282,7 +283,8 @@ with common.get_session(options.gpus) as session:
                 point_sizes=np.array(list(point_sizes)),
                 perplexity=options.tsne_perplexity,
                 save_to=options.figure_dir,
-                subject='MOA')
+                subject='MOA',
+                label_names=moa_names)
 
         if options.vector_distance:
             try:

### 2017-10-18 cytogan/train/visualize.py

@@ -64,7 +64,8 @@ def latent_space(latent_vectors,
                  perplexity=None,
                  point_sizes=None,
                  save_to=None,
-                 subject=None):
+                 subject=None,
+                 label_names=None):
     assert np.ndim(latent_vectors) == 2
     log.info('Plotting latent space for %d vectors', len(latent_vectors))
 
@@ -73,6 +74,7 @@ def latent_space(latent_vectors,
     if isinstance(perplexity, int):
         perplexity = [perplexity]
 
+    log.info('Computing TSNEs at perplexity %s', tuple(perplexity))
     for p in perplexity:
         reduction = sklearn.manifold.TSNE(
             n_components=2, perplexity=p, init='pca', verbose=1)
@@ -88,6 +90,10 @@ def latent_space(latent_vectors,
             c=labels,
             lw=point_sizes,
             cmap=plot.cm.Spectral)
+
+        if label_names is not None:
+            colorbar = plot.colorbar()
+            colorbar.ax.set_yticklabels(label_names)
 
         if save_to is not None:
             subject_suffix = '-{0}'.format(subject.lower()) if subject else ''

### 2017-10-18 cytogan/metrics/profiling.py
Fixed visualizations

@@ -77,8 +77,8 @@ def log_top_k(compound, distances, test_data, training_data, k=3):
         for s, j in zip(top_k_moas, y):
             top_k_strings.append('{0} ({1:.8f})'.format(s, distances[i, j]))
         top_k_string = ', '.join(top_k_strings)
-        log.info('Top %d neighbors for %s/%.4f (%s): %s'.format(
-            k, compound, c, m, top_k_string))
+        log.info('Top %d neighbors for %s/%.4f (%s): %s', k, compound, c, m,
+                 top_k_string)
 
 
 def get_nearest_neighbors(examples, neighbors):
@@ -129,8 +129,7 @@ def score_profiles(dataset):
         log.info('Accuracy for %s is %.3f', holdout_compound, accuracy)
         accuracies.append(accuracy)
 
-        for actual in actual_labels:
-            for predicted in predicted_labels:
-                confusion_matrix[actual][predicted] += 1
+        for actual, predicted in zip(actual_labels, predicted_labels):
+            confusion_matrix[actual][predicted] += 1
 
     return confusion_matrix, np.mean(accuracies)

### 2017-10-18 cytogan/train/visualize.py

@@ -92,7 +92,8 @@ def latent_space(latent_vectors,
             cmap=plot.cm.Spectral)
 
         if label_names is not None:
-            colorbar = plot.colorbar()
+            ticks = list(range(len(label_names)))
+            colorbar = plot.colorbar(ticks=ticks)
             colorbar.ax.set_yticklabels(label_names)
 
         if save_to is not None:
@@ -208,7 +209,7 @@ def confusion_matrix(matrix,
                      title='Confusion Matrix',
                      accuracy=None,
                      save_to=None):
-    figure, axis = plot.subplots(figsize=(14, 10))
+    figure, axis = plot.subplots(figsize=(14, 12))
     if accuracy:
         title += ' ({0:.1f}% Accuracy)'.format(accuracy * 100)
     figure.suptitle(title)

### 2017-11-15 cytogan/experiments/visualize.py
Generalized option to save generated images for bbbc021

@@ -393,6 +393,14 @@ def image_algebra(model,
         if save_to is not None:
             _save_figure(save_to, 'image-algebra-vectors.png')
 
+def save_images(images, directory):
+    if not os.path.exists(directory):
+        os.makedirs(directory)
+    for n, image in enumerate(images):
+        path = os.path.join(directory, '{0}.png'.format(n))
+        scipy.misc.imsave(path, image)
+    log.info('Saved %d images to %s', len(images), directory)
+
 
 def disable_display():
     plot.switch_backend('Agg')

### 2017-11-15 cytogan/train/bbbc021.py

@@ -42,6 +42,7 @@ parser.add_argument('--no-latent-embedding', action='store_true')
 parser.add_argument('--noise-file')
 parser.add_argument('--normalize-luminance', action='store_true')
 parser.add_argument('--save-profiles', action='store_true')
+parser.add_argument('--save-generated-images', action='store_true')
 parser.add_argument('--skip-evaluation', action='store_true')
 parser.add_argument('--store-generated-noise', action='store_true')
 parser.add_argument('--tsne-perplexity', type=int)
@@ -477,14 +478,10 @@ with common.get_session(options.gpus, options.random_seed) as session:
             if np.ndim(noise) == 1:
                 noise = noise.reshape(1, -1)
             images = model.generate(noise)
-            directory = os.path.join(options.workspace, options.figure_dir,
-                                     'from-noise')
-            if not os.path.exists(directory):
-                os.makedirs(directory)
-            for n, image in enumerate(images):
-                path = os.path.join(directory, '{0}.png'.format(n))
-                scipy.misc.imsave(path, image)
-                log.info('Saved image generated from noise: %s', path)
+
+        if options.save_generated_images:
+            directory = os.path.join(options.figure_dir, 'generated')
+            visualize.save_images(images, directory)
 
 if options.show_figures:
     visualize.show()

### 2017-11-15 cytogan/experiments/visualize.py
visualize.generative_samples returns images now

@@ -236,6 +236,8 @@ def generative_samples(model,
     if save_to is not None:
         _save_figure(save_to, filename)
 
+    return images
+
 
 def confusion_matrix(matrix,
                      title='Confusion Matrix',

### 2017-11-15 cytogan/train/bbbc021.py

@@ -456,7 +456,7 @@ with common.get_session(options.gpus, options.random_seed) as session:
             labels = labels.repeat(options.generative_samples // 2, axis=0)
             samples.append(labels)
 
-        visualize.generative_samples(
+        images = visualize.generative_samples(
             model,
             samples,
             save_to=options.figure_dir,

### 2017-11-15 paper/paper.tex
Updated paper a bit

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[final]{nips_2017}
-% \usepackage{nips_2017}
+% \usepackage[final]{nips_2017}
+\usepackage{nips_2017}
 
@@
 
+\graphicspath{{figures/}}
+
 %\PassOptionsToPackage{numbers}{natbib}
@@
 
-\title{Generative Modeling of Cell Images}
+\title{CytoGAN: Generative Modeling of Cell Images}
+% * <anne@broadinstitute.org> 2017-10-15T19:01:17.646Z:
+%
+% > CytoGAN
+% Is it necessary to give it a name? This seems overly common to me, but then again maybe it's the convention and you just go with it.
+%
+% ^.
 
@@
 When evaluated for their ability to group cell images responding to treatment by chemicals of known classes, we find that adversarially learned representations are superior to autoencoder-based approaches.
-While currently inferior to classical computer vision and transfer learning, the adversarial framework enables useful and previously impossible interpretability of latent codes.
+While currently inferior to classical computer vision and transfer learning, the adversarial framework enables useful visualization of the variation of cellular images due to their generative capabilities.
 \end{abstract}
@@
 \section{Introduction}
-Advances in high-throughput microscopy systems have enabled acquisition of large
-volumes of high-resolution cell images \cite{Kraus2016}. This paves the way for
-novel computational methods that can leverage these large quantities of data to
-study biological systems. Our work focuses on the task of morphological
-profiling, which aims to map microscopy images of cells to representation
-vectors that capture salient features and axes of variation in an unsupervised
-manner \cite{caicedo_profiling}. These representations ideally divide the
-morphological space into clusters of cells with similar properties or, in the
-case of induced perturbations, similar function.
+Advances in high-throughput microscopy systems have enabled acquisition of large volumes of high-resolution cell images \cite{caicedo2017data}.
+This paves the way for novel computational methods that can leverage these large quantities of data to study biological systems.
+Our work focuses on the task of morphological profiling, which aims to map microscopy images of cells to representation vectors that capture salient features and axes of variation in an unsupervised manner \cite{caicedo_profiling}.
+These representations ideally divide the morphological space into clusters of cells with similar properties or, in the case of induced perturbations, similar function.
 
-Current techniques for morphological profiling broadly fall in two categories:
-a) Classical image processing, using specialized software like CellProfiler
-\cite{Carpenter2006}, to capture representations via manually-tuned segmentation
-and traditional computer vision pipelines, and b) transfer learning to extract
-features learned by deep convolutional neural networks originally trained to
-classify miscellaneous objects \cite{pawlowski2016automating,
-ando2017improving}. Classical computer vision approaches offer better
-interpretability of features but require more human tuning of the segmentation
-algorithms, and are limited by the feature set implemented in the image analysis
-software. Current transfer learning approaches have been shown to outperform
-classical methods in at least one dataset \cite{ando2017improving}. However,
-given that these networks were trained on natural images (in RGB), they do not
-discover the relations of the biologically meaningful image channels. Instead,
-their superior performance is likely due to the their ability to capture
-high-level vision features, which represent the overall structure of cells, but
-not necessarily intricate details of their morphological variations. Therefore,
-we hypothesized that learning representations specifically tuned to cell images
-would be valuable.
+Current techniques for morphological profiling broadly fall in two categories: a) classical image processing, using specialized software like CellProfiler \cite{Carpenter2006} to capture representations via manually-tuned segmentation and traditional computer vision pipelines, and b) transfer learning to extract features learned by deep convolutional neural networks originally trained to classify miscellaneous objects \cite{pawlowski2016automating, ando2017improving}.
+Classical computer vision approaches offer better interpretability of features but require more human tuning of the segmentation algorithms, and are limited by the feature set implemented in the image analysis software.
+Current transfer learning techniques have been shown to outperform classical methods in at least one dataset \cite{pawlowski2016automating,ando2017improving}.
+However, given that these networks were trained on natural images (in RGB), they do not discover the relations of the biologically meaningful image channels.
+Instead, their superior performance is likely due to their ability to extract high-level vision features, which appear to capture the overall structure of cells, but not necessarily all the intricate details of their morphological variations.
+Therefore, we hypothesized that learning representations specifically adapted to cell images would be valuable.
 
-We employ Generative Adversarial Networks (GANs) \cite{goodfellow2014generative}
-to build a generative model of cell images from the BBBC021 dataset
-\cite{Caie2010,Ljosa2012}.  We show that this generative model learns rich
-feature representations and synthesizes realistic images with interpretable
-properties. Our approach offers the following advantages:
-
+We employ Generative Adversarial Networks (GANs) \cite{goodfellow2014generative} to build a generative model of single cell images from the BBBC021 dataset \cite{Ljosa2012}.
+We show that this model learns rich feature representations and synthesizes realistic images that are useful for exploring morphological variation in cells.
+The main advantages of this method are:
 \begin{itemize}
-  \item \emph{Adaptability}: Unsupervised representation learning, which enables easy adaptation and incorporation of new data.
-This is important given that applications of morphological profiling do not have ground truth annotations, but instead aim to reveal the structure of the image collection.
-  \item \emph{Specialization}: Our method learns from training data and is able to extract inherent semantic relationships.
-Transfer learning approaches lack this ability and cannot capture the intrinsic relationships between the biologically meaningful channels.
-  \item \emph{Interpretability}: The generative abilities of this model enable previously impossible interpretbility by being able to visualize cells dependent on a given representation.
+  \item \emph{Adaptability}: Unsupervised representation learning, which can readily incorporate new data.
+This is important given that applications of morphological profiling usually lack ground truth annotations, but instead aim to reveal the structure of the image collection.
+  \item \emph{Specialization}: Because this method learns from its training data it is able to disentangle factors of variation and extract inherent semantic relationships.
+Transfer learning approaches lack this ability and cannot capture the intrinsic relations between the biologically meaningful channels.
+  \item \emph{Translating into biological phenotypes}: The generative abilities of this model enable useful visualizations of cells to help translate data variations into meaningful biological phenotypes.
 \end{itemize}
@@
 
-Our work lies at the intersection of automated morphological profiling, and
-representation learning with deep neural networks, in particular generative
-architectures. Caicedo et al. \cite{caicedo2017data} recently outlined the state
-and challenges of the morphological profiling problem. Prior to this, Ljosa et
-al. \cite{Ljosa2013} compared the performance of various dimensionality
-reduction methods for CellProfiler features on the BBBC021 benchmark
-\cite{Caie2010,Ljosa2012}, consisting of MCF7 cells exposed to different
-chemical treatments. Singh et al. \cite{Singh2014} demonstrated the importance
-of illumination correction of those images. Pawlowski et al.
-\cite{pawlowski2016automating} for the first time reported a
-representation-learning method based on deep learning that is competitive with
-hand-engineered features at the task of MOA prediction. Ando et al.
-\cite{ando2017improving} also applied transfer learning with a different
-architecture and introducing a novel feature normalization method. Further
-related work on this dataset spans supervised classification \cite{Kraus2016a}
-or deep learning applications based on CellProfiler features
-\cite{Kandaswamy2016,Zamparo2015}
+Our work lies at the intersection of automated morphological profiling, and representation learning with deep neural networks, in particular generative architectures.
+Caicedo et al.
+\cite{caicedo2017data} recently outlined the state and challenges of the morphological profiling problem.
+Prior to this, Ljosa et al.
+\cite{Ljosa2013} compared the performance of various dimensionality reduction techniques for CellProfiler features on the BBBC021 benchmark \cite{Ljosa2012}, consisting of MCF7 cells exposed to different chemical treatments.
+Pawlowski et al.
+\cite{pawlowski2016automating} for the first time reported a representation-learning method based on deep learning that is competitive with hand-engineered features at the task of predicting mechanism-of-action (MOA) of chemicals.
+Ando et al.
+\cite{ando2017improving} also applied transfer learning with a different architecture and introduced a novel feature normalization method.
+Other related work on this dataset include supervised classification \cite{Kraus2016a}, transfer learning on CellProfiler features \cite{Kandaswamy2016}, and dimensionality reduction using autoencoders \cite{Zamparo2015}.
 
-Few works have applied unsupervised deep learning techniques to the task of
-feature extraction in morphological profiling. Pawlowski \cite{pawlowski2016msc}
-first investigated autoencoder-based approaches but reported results far
-inferior to hand-tuned features and transfer-learning approaches. Our model is
-more related to the work by Osokin et al. \cite{osokin2017gans} and Johnson et
-al. \cite{johnson2017generative}, wherein GANs were used to model cell
-images, although their applications did not include morphological profiling.
+Few published studies have applied unsupervised deep learning techniques to the task of feature extraction in morphological profiling.
+Pawlowski \cite{pawlowski2016msc} first investigated autoencoder-based methods but reported results far inferior to hand-tuned features or transfer-learning approaches.
+Our model is more related to the work by Osokin et al.
+\cite{osokin2017gans} and Johnson et al.
+\cite{johnson2017generative}, wherein GANs were used to model cell images, although their applications did not include morphological profiling.
 
-\section{Using GANs for Interpretable Representation Learning}
+\section{Using GANs for Representation Learning}
+Goodfellow et al. \cite{goodfellow2014generative} introduced GANs as a game of
+two players: a \emph{generator} $G$ and a \emph{discriminator} or \emph{critic} $D$.
+The former receives samples $\mathbf{z}$ drawn from a noise prior $P_{\text{noise}}$ which it maps to values $G(\mathbf{z})$ that should resemble elements of some data distribution $P_{\text{data}}$.
+The discriminator must learn to distinguish such synthetic samples from real values $\mathbf{x} \sim P_{\text{data}}$.
+The critic's confidence in the realism of the generator's productions is used as feedback to $G$, guiding it to synthesize ever more realistic replicates of samples from the data prior.
+This procedure is formalized in a zero-sum game,
+\vspace{-1pt}
+$$\min_G \max_D V(G, D) = \mathbb{E}_{\mathbf{x} \sim P_{\text{data}}}[\log(D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim P_{\text{noise}}}[\log(1 - D(G(\mathbf{z}))].$$
+%\vspace{-1pt}
+Radford et al.
+\cite{radford2015unsupervised} first specialized GANs to image synthesis by introducing Deep Convolutional GANs (DCGANs).
+DCGANs implement the generator and discriminator based on convolutional operations.
+Derivations of DCGANs such as Least Squares GAN (LSGAN) \cite{mao2016least} or Wasserstein GAN (WGAN) \cite{arjovsky2017wasserstein} tackle instabilities in the training procedure of early generative adversarial models.
+In our experiments, we found LSGAN to be most stable, in part leading to higher quality generated images than both DCGAN and WGAN.
 
-Goodfellow et al. \cite{goodfellow2014generative} introduced GANs as a game of
-two \emph{players}: a \emph{generator} $G$ and a \emph{discriminator} $D$. The
-former receives samples $\mathbf{z}$ drawn from a noise prior $P(\mathbf{z})$
-which it maps to values $G(z)$ that should resemble elements of some data
-distribution $P_{\text{data}}$. The discriminator must learn to distinguish such
-synthetic samples from real values $\mathbf{x} \sim
-P_{\text{data}}(\mathbf{x})$. The critic's confidence in the realism of the
-generator's productions is used as feedback to $G$, guiding it to synthesize
-ever more realistic replicates of samples from the data prior. This procedure is
-formalized in a zero-sum game, solved in a nash equilibrium found by playing the
-minimax game
+The original GAN framework does not include an explicit means of performing inference.
+As such, we require extensions that allow mapping of a sample $\mathbf{x}$ drawn from the data prior to a latent representation via some encoder transformation $E(\mathbf{x})$.
+A common approach is to interpret the penultimate layer of the discriminator as a latent space.
+The activations of this layer serve as a source of representation vectors.
+These latent codes are thought to be meaningful because the discriminator must develop a strong internal representation of its input to succeed at its discrimination task.
+Furthermore, this method imposes no computational overhead compared to vanilla GANs.
+More advanced encoders could tie this encoding to a corresponding noise vector.
 
-$$\min_G \max_D V(G, D) = \mathbb{E}_{x \sim P_{data}}[\log(D(\mathbf{x})] + \mathbb{E}_{z \sim P_{noise}}[\log(1 - D(G(\mathbf{z}))].$$
+\section{Exploring Biological Phenotypes Using Cell Image Synthesis}\label{synthesis}
+% Exploring Biological Phenotypes Using Cell Image Synthesis
+GANs have been proven to synthesize realistic images both within and beyond the biological domain \cite{goodfellow2014generative,radford2015unsupervised,osokin2017gans}.
+We examine if this ability transfers to cell images extracted from the BBBC021 dataset of human breast cancer cell lines.
+Each image in BBBC021 consists of three \emph{channels} corresponding to DNA, F-Actin and $\beta$-Tubulin, cellular components visualized by fluorescence microscopy.
+We stack these channels and treat them as RGB images via a simple $\text{DNA} \mapsto \text{R}, \text{$\beta$-Tubulin} \mapsto \text{G}, \text{F-Actin} \mapsto \text{B}$ mapping.
+We obtain ca.
+1.3 million single cell images by segmenting the raw images using CellProfiler.
+Finally, we normalize the luminance of each channel to the range [0 -- 1].
 
-Radford et al. \cite{radford2015unsupervised} first specialized GANs to image
-synthesis by introducing Deep Convolutional GANs (DCGANs). DCGANs implement the
-generator and discriminator based on convolutional operations. Derivations of
-DCGANs such as Least Squares GAN (LSGAN) \cite{mao2016least} or Wasserstein GAN
-(WGAN) \cite{arjovsky2017wasserstein} tackle infamous instabilities in the
-training procedure of early generative models. In our experiments, we found
-LSGAN to be most stable, in part leading to higher quality generated images than
-both DCGAN and WGAN.
-
-The original GAN framework does not include an explicit means of performing
-inference. As such, we require extensions that allow mapping of a sample
-$\mathbf{x}$ drawn from the data prior to a latent representation via some
-encoder transformation $E(\mathbf{x})$. A common approach is to interpret the
-penultimate layer of the discriminator as latent space. The activations of this
-layer serve as a source of representation vectors. These latent codes are
-thought to be meaningful because the discriminator must develop a strong
-internal representation of its input in order to succeed at its discrimination
-task. Furthermore, this method imposes no computational overhead compared to the
-vanilla GAN models.
-
-\section{Interpretable Synthesis of Cells}\label{synthesis}
-
-We examine if GAN models' reported ability to synthesize realistic images
-transfers to the domain of cell microscopy. Our data prior consists of images
-from the BBBC021 dataset of human breast cancer cells. Originally, each image
-consists of three \emph{channels} corresponding to DNA, F-Actin and B-Tubulin,
-cellular components visualized by fluorescence microscopy. We stack these
-channels and treat them as RGB images via a simple $\text{DNA} \mapsto \text{R},
-\text{$\beta$-Tubulin} \mapsto \text{G}, \text{F-Actin} \mapsto \text{B}$
-mapping. We obtain ca. 1.3 million single cell images by segmenting the raw
-images using CellProfiler software. Finally, we normalize the luminance of each
-channel to the range [0 -- 1].
+% Current approaches to morphological profiling extract features that are challenging to translate into biological meaning, such as Zernike moments.
+While other features capture readily understandable concepts such as the area occupied by the nucleus, even these are difficult to interpret when a given class of cells is defined by several such features in combination.
+For transfer learning it is nearly impossible to visualize the concept of a feature.
+GANs present a solution as they hold a generative model that enables the visualization of the impact of perturbations along a feature of the noise space.
+Further, more advanced GANs integrate an inference method to make the feature and noise space identical such that visualization of feature vectors becomes possible.
+%this issue by ,  we found that GAN representations are not yet competitive with those extracted by other approaches, we believe a major benefit provided by generative adversarial models is their interpretability.
+We showcase this in the following.
 
@@
 realistic, but also consistent with their biological nature. For example, it is
-characteristic that B-Tubulin forms a circular halo cradling the nucleus of a
-cell. This property is maintained clearly in all generated images.
+characteristic that $\beta$-Tubulin forms a circular halo cradling the nucleus.
+This property is maintained clearly in all generated images.
 
@@
   \begin{tikzpicture}
-    \tikzset{frame/.style={draw, semithick, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt}}
-
   	\draw (0, 0) node {\includegraphics[width=0.6\textwidth]{generated}};
 
-    \path (-3.3, 0) coordinate [frame] (r);
+    \path (-3.3, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (r);
     \draw (r)+(0, -2.2) node {\textbf{Real}};
 
-    \path (-1.07, 0) coordinate [frame] (l);
+    \path (-1.07, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (l);
     \draw (l)+(0, -2.2) node {\textbf{LSGAN}};
 
-    \path (1.14, 0) coordinate [frame] (w);
+    \path (1.14, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (w);
     \draw (w)+(0, -2.2) node {\textbf{WGAN}};
 
-    \path (3.35, 0) coordinate [frame] (d);
+    \path (3.35, 0) coordinate [draw, rectangle, text width=1.7cm, text height=3.5cm, rounded corners=3pt] (d);
     \draw (d)+(0, -2.2) node {\textbf{DCGAN}};
   \end{tikzpicture}
-  \caption{Examples of real MCF7 cells from the BBBC021 dataset juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN.
-The generated images of LSGAN are most realistic and consistent with biological properties of cells.}
+  \caption{Examples of real BBBC021 cell images juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN.
+LSGAN synthesizes the most realistic images as judged by our expert biologists.}
   \label{fig:generated}
@@
 
-Current approaches to morphological profiling capture representations that are
-often hard to interpret. While certain CellProfiler features such as nucleic
-area are easy to understand, others like Zernike moments are not intuitive. For
-transfer learning it is nearly impossible to visualize the captured concept. In
-contrast, the GAN noise space $P(z)$ has been shown to be highly interpretable
-and reveal rich semantic structure \cite{radford2015unsupervised}. We are able
-to demonstrate this for images of cells. Figure \ref{fig:interpolations}
-exhibits how interpolating between two noise samples $z_1, z_2 \sim P(z)$ leads
-to visually smooth transitions in synthesized cells. Figure \ref{fig:algebra}
-shows that $P(z)$ encodes semantic relationships, enabling algebra on
-interpretable properties of the images. While there is no way to encode images
-into $P(\mathbf{z})$ with the methods presented so far, we believe more advanced
-models that enable this will be highly valuable if they can maintain these rich
+Current approaches to morphological profiling extract features that are
+challenging to translate into biological meaning, such as Zernike moments. While
+other features capture readily understandable concepts such as the area occupied
+by the nucleus, even these are difficult to interpret when a given class of
+cells is defined by several such features in combination. For transfer learning
+it is nearly impossible to visualize the concept of a feature. In contrast, the
+noise space of GANs is known to be highly interpretable and reveal rich
+semantic structure \cite{radford2015unsupervised}. We are able to demonstrate
+this for images of cells. Figure \ref{fig:interpolations} exhibits how
+interpolating between two noise samples leads to smooth transitions in synthesized cells.
+This supports the fact that the generator learned a low-dimensional manifold of the images.
+Figure \ref{fig:algebra} shows
+that the noise space encodes semantic relationships, enabling algebra on
+interpretable properties of the images such as size of the nucleus or $\beta$-Tubulin content.
+While there is no way to encode images into $P_{\text{noise}}$ with the methods presented so far, we believe more advanced
+architectures that enable this will be highly valuable if they can maintain these remarkable
 semantic properties.
@@
   \includegraphics[width=\textwidth, trim={0 0 0 2.5cm}, clip]{interpolation}
-  \caption{Interpolating between two points $\mathbf{z_1}, \mathbf{z_2} \sim P(\mathbf{z})$ results in visually smooth transitions in the corresponding synthesized images.
-Each row shows the transition from $\mathbf{z_1}$ to $\mathbf{z_2}$ from left to right.}
-  \vspace{-0.1cm}
+  \caption{Interpolating between two points $\mathbf{z}_1, \mathbf{z}_2 \sim P_{\text{noise}}$ results in smooth transitions in the synthesized cell images.
+Each row shows the transition from $\mathbf{z}_1$ to $\mathbf{z}_2$ from left to right.}
+  \label{fig:interpolations}
 \end{figure}
@@
   \caption{Vector algebra in the noise space translates to biologically valuable relationships in generated images.
-In the first row, subtracting the vector for a cell with small quantities of $\beta$-Tubulin (green stain) from one with high amounts yields a vector representation for "high $\beta$-Tubulin content".
+In the first row, subtracting the vector for a cell with small quantities of $\beta$-Tubulin (green stain) from one with high amounts yields a vector representation for "higher $\beta$-Tubulin content".
 In the second row, the difference between a large and small nucleus encodes the semantic meaning of "larger nucleus", which can be added to vectors of other cell images to grow the size of the nucleus.}
@@
 \end{figure}
+% We show the generative abilities of our models and the richness of the noise space on three ways: a) The model generates high quality cell images as shown in Figure \ref{fig:generated}.
+The LSGAN produces the most convincing images.
+b) The noise space is capable of smooth interpolations between two noise samples $z_1, z_2 \sim P(z)$, displayed in Figure \ref{fig:interpolations}.
+This supports the fact that the generator learned low-dimensional manifold of the images.
+c) $P(z)$ encodes semantic information that enables algebra based on semantically relevant properties of the images.
+Figure \ref{fig:algebra} presents examples of vector algebra in $P(z)$ with biologically interpretable results.
+We are able to leverage the linear relationships present in the latent space to influence the size of synthesized nuclei as well as $\beta$-Tubulin content.
 
 \section{Representation Learning for Morphological Profiling}\label{moa}
-We test the quality of representations extracted by the discriminator used to generate single cell images learned by testing their ability to cluster treatments of similar function or mechanism-of-action.
+We test the quality of representations extracted via the discriminator by evaluating their ability to cluster treatments of similar function, here the mechanism-of-action.
 We obtain treatment profiles by averaging the extracted single cell representations found as intermediate activations of the discriminator.
-Further, we follow the experimental protocol of \cite{Ljosa2013} and report an average mechanism-of-action classification accuracy of a leave-one-compound-out cross-validation using a one nearest neighbor classifier.
-
+Further, we follow the experimental protocol of \cite{Ljosa2013} and report an average MOA classification accuracy of a leave-one-compound-out cross-validation using a one nearest neighbor classifier.
 \begin{table}
@@
   \end{tabular}
-  \caption{Accuracy of various methods at the task of predicting the mechanism-of-action of treatments via nearest-neighbor classification in the BBBC021 dataset.
+  \caption{Comparison of the mechanism-of-action classification accuracy.
 CP refers to the results by Singh et al.
-\cite{Singh2014} using CellProfiler Features.
+\cite{Singh2014} using CellProfiler features.
 The Deep Transfer methods correspond to Deep Feature Transfer as proposed by Pawlowski et al.
@@
   \label{moa-results}
+  \vspace{-0.5cm}
 \end{table}
 
-Table \ref{moa-results} compares the accuracy of our approach to classical CellProfiler features and transfer learning.
-We found that the quality of representations found by unsupervised learning is not yet competitive with other established methods.
-Nevertheless, we believe that the adversarial approach has significant benefits regarding its interpretability as outlined above.
-Additionally, this framework is able to adapt to the dataset at hand and extract inherent relations that are not captured by previous methods.
-We believe that further gains can be made, as the image quality corresponds to classification accuracy.
-As such the best performing method (LSGAN) also yields the highest image quality, as judged qualitatively by expert biologists and shown in Figure \ref{fig:generated}.
+Table \ref{moa-results} compares the accuracy of our technique to classical CellProfiler features and transfer learning.
+We found that the quality of representations extracted by our approach is not yet competitive with other methods.
+Nevertheless, we believe that the adversarial approach has significant benefits regarding its ability to translate into biological phenotypes as outlined above.
+Additionally, this framework is able to adapt to the dataset at hand and extract inherent relations that are not captured by previous techniques.
+We believe that further improvements are possible, as the quality of synthesized images correlates with classification accuracy.
+Accordingly, the best performing method (LSGAN) also yields the highest image quality, as judged qualitatively by expert biologists and shown in Figure \ref{fig:generated}.
 
 \section{Conclusion}
+This work investigates the use of GANs for the domain of cell microscopy imaging, in particular morphological profiling.
+First, we demonstrate the abilities of our generative model with the exploration of morphological phenotypes by synthesizing realistic images of cells and performing transformations on them such as interpolation and vector algebra.
+Second, we extend standard GANs with an encoder and assess the quality of learned representations by evaluating their mechanism-of-action classification performance following \cite{Ljosa2013}.
+Even though adversarially learned representations are currently inferior at this task, we argue that further enhancements to GANs and encoding schemes will lead to biologically richer latent representations and better MOA classification accuracy.
 
-This work investigates the use of GANs for the domain of cell microscopy
-imaging, particularly morphological profiling. First, we demonstrate that GANs
-are able to synthesize biologically realistic images of cells. In addition, we
-explain that GANs induce a rich semantic structure in their noise space,
-providing a powerful basis for interpretability of observed properties. Second,
-we add an encoding component to standard GANs and evaluate the quality of
-learned representations via their mechanism-of-action classification performance
-following \cite{Ljosa2013}. Even though adversarially learned representations
-are currently inferior at this task, we argue that further enhancements to GANs,
-especially biologically meaningful ones, will lead to higher quality latent
-representations.
+This work covers only a small fraction of possible applications of GANs to the domain of microscopy images.
+For example, we hope that future work investigates BiGANs\cite{donahue2016adversarial} or other novel solutions to infer latent features with GANs.
+We believe that improved inference, combined with the interpretable nature of the GAN framework, may enable simulated experiments by performing algebra with vectors corresponding to cell lines, diseases or other perturbations.
+Finally, we note that deep learning generally flourishes with larger amounts of data.
+Thus, we conjecture that the performance of the outlined approaches will improve with larger datasets.
+% forcing new page so that references are always on Page 5.
+\newpage
 
-We emphasize that this work covers only a very small fraction of possible
-applications of GANs to the domain of microscopical images. For example, we hope
-that future work investigates BiGANs \cite{donahue2016adversarial} or other
-novel solutions to infer latent features with GANs. We believe that improved
-inference, combined with the interpretable nature of the GAN framework, may
-enable simulated experiments by performing algebra with vectors corresponding to
-cell lines, diseases or other perturbations. Finally, we note that deep learning
-generally flourishes with larger amounts of data. As such, we conjecture that
-the performance of the outlined approaches will improve with the use of larger
-datasets.
-
-\section*{Acknowledgements}
-Nick Pawlowski is supported by Microsoft Research through its PhD Scholarship
-Program and the EPSRC Centre for Doctoral Training in High Performance
-Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
+%\section*{Acknowledgements}
+%NP is supported by Microsoft Research through its PhD Scholarship Programme and the EPSRC Centre for Doctoral Training in High Performance Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
 

### 2017-11-15 scripts/normalize_luminance.py

@@ -31,7 +31,10 @@ try:
     for path in tqdm.tqdm(options.images):
         if options.folder:
             path = os.path.join(options.folder, path)
-        image = scipy.misc.imread(path).astype(np.float32) / 255.0
+        try:
+            image = scipy.misc.imread(path).astype(np.float32) / 255.0
+        except:
+            continue
         normalized = normalize_luminance(image)
         out_path = os.path.join(options.out, os.path.basename(path))
         scipy.misc.imsave(out_path, normalized)

### 2017-11-23 paper/paper.tex
Updated paper

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{nips_2017}
-\usepackage{nips_2017}
+\usepackage[final]{nips_2017}
+% \usepackage{nips_2017}
 
+\usepackage[dvipsnames]{xcolor}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 
+\usepackage{tikz}
+\usetikzlibrary{3d}
+\usetikzlibrary{shapes}
+\usepgflibrary{fpu}
+
 \graphicspath{{figures/}}
 
-%\PassOptionsToPackage{numbers}{natbib}
-%\usepackage[numbers]{natbib}
-%\usepackage[natbib, maxcitenames=3, style=numeric]{biblatex}
-%\setcitestyle{numbers}
-%\bibliographystyle{abbrvnat}
+% \PassOptionsToPackage{numbers}{natbib}
+% \usepackage[numbers]{natbib}
+% \usepackage[natbib, maxcitenames=3, style=numeric]{biblatex}
+% \setcitestyle{numbers}
+% \bibliographystyle{abbrvnat}
 
 \title{CytoGAN: Generative Modeling of Cell Images}
-% * <anne@broadinstitute.org> 2017-10-15T19:01:17.646Z:
-%
-% > CytoGAN
-% Is it necessary to give it a name? This seems overly common to me, but then again maybe it's the convention and you just go with it.
-%
-% ^.
 
@@
   Peter Goldsborough\thanks{This work has been conducted as part of an internship at the Imaging Platform of the Broad Institute.}\\
-  %Imaging Platform \\
+  Imaging Platform \\
   Broad Institute of MIT and Harvard \\
-  %Cambridge, MA, USA \\
+  Cambridge, MA, USA \\
   \texttt{pgoldsbo@broadinstitute.org} \\
@@
   Nick Pawlowski \\
-  %Department of Computing \\
+  Department of Computing \\
   Imperial College London \\
-  %London, UK \\
+  London, UK \\
   \texttt{n.pawlowski16@imperial.ac.uk} \\
@@
   Juan C Caicedo \& Shantanu Singh \& Anne E Carpenter\\
-  %Imaging Platform \\
+  Imaging Platform \\
   Broad Institute of MIT and Harvard \\
-  %Cambridge, MA, USA \\
+  Cambridge, MA, USA \\
   \texttt{\{jccaicedo, shsingh, anne\}@broadinstitute.org} \\
@@
 $$\min_G \max_D V(G, D) = \mathbb{E}_{\mathbf{x} \sim P_{\text{data}}}[\log(D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim P_{\text{noise}}}[\log(1 - D(G(\mathbf{z}))].$$
-%\vspace{-1pt}
+\vspace{-1pt}
 Radford et al.
@@
 Furthermore, this method imposes no computational overhead compared to vanilla GANs.
+Figure \ref{fig:gan} shows a simplified schematic of our GAN architecture.
 More advanced encoders could tie this encoding to a corresponding noise vector.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[thick]
+    \tikzset{box/.style={
+      draw, rectangle, rounded corners=0.8pt,
+      text width=2.2cm, text height=0.75cm}}
+    \tikzset{node/.style={draw, circle, inner sep=0.35cm}}
+
+    % Model
+    \path (0, 0) coordinate [draw, circle, inner sep=5pt]
+          (z) node {$\mathbf{z}$};
+
+    \path (2, 0) coordinate [box] (G) node {Generator};
+
+    \path (4.3, -0.8) coordinate [node] (f) node {$\mathbf{x}$};
+    \path (4.3, +0.8) coordinate [node] (r) node {$G(\mathbf{z})$};
+
+    \path (9.5, 0) coordinate [draw, circle, inner sep=7pt]
+          (D) node {$\hspace{1pt}\mathbf{D}$};
+
+    % Anchor for real and fake arrows.
+    \path (5, 0) coordinate (_);
+
+    \foreach \x/\l in {0/Convolution,%
+                       1/Convolution,%
+                       2/Convolution,%
+                       3/{\hspace{1mm}Dense(\small$\ell$)\textsuperscript{\color{Red}\textdagger}},%
+                       4/{Dense({\small$1$})}%
+                       } {
+      \path ({5.8 + \x * 0.65}, 0)
+            coordinate [box, text width=0.3cm, text height=2cm] (l\x)
+            node [rotate=90] {\l};
+    }
+
+    % Footnote
+    \draw (7, -1.5)
+          node {\footnotesize{\color{Red}\textdagger} Representations};
+
+    % Edges
+    \draw (f) -- (_);
+    \draw (r) -- (_);
+    \draw [->] (_) -- (l0);
+    \draw [->] (l4) -- (D);
+    \draw [->] (z) -- (G);
+    \draw [->] (G) edge[out=0] (f);
+  \end{tikzpicture}
+  \caption{A simplified diagram of the modified DCGAN \cite{radford2015unsupervised} architecture we employ for our experiments.
+On a high level, the discriminator is a sequence of convolutional layers, leading to a fully connected (dense) layer with $\ell$ neurons.
+We interpret the activations of these $\ell$ neurons as latent vectors and further feed them into a final dense layer with just one activation, forming the discriminator's output.
+}
+  \label{fig:gan}
+\end{figure}
 
@@
 
-% Current approaches to morphological profiling extract features that are challenging to translate into biological meaning, such as Zernike moments.
-While other features capture readily understandable concepts such as the area occupied by the nucleus, even these are difficult to interpret when a given class of cells is defined by several such features in combination.
-For transfer learning it is nearly impossible to visualize the concept of a feature.
-GANs present a solution as they hold a generative model that enables the visualization of the impact of perturbations along a feature of the noise space.
-Further, more advanced GANs integrate an inference method to make the feature and noise space identical such that visualization of feature vectors becomes possible.
-%this issue by ,  we found that GAN representations are not yet competitive with those extracted by other approaches, we believe a major benefit provided by generative adversarial models is their interpretability.
-We showcase this in the following.
-
 Figure \ref{fig:generated} shows examples of images generated with LSGAN, WGAN
@@
 
-\begin{figure}
+\begin{figure}[b]
   \centering
@@
     \midrule
-    43 \%& 45 \% & 63 \% & 49 \% & 90 \% & 91 \% & 96 \% \\
+    51 \%& 56 \% & 68 \% & 49 \% & 90 \% & 91 \% & 96 \% \\
     \bottomrule \\
@@
 Thus, we conjecture that the performance of the outlined approaches will improve with larger datasets.
-% forcing new page so that references are always on Page 5.
-\newpage
 
-%\section*{Acknowledgements}
-%NP is supported by Microsoft Research through its PhD Scholarship Programme and the EPSRC Centre for Doctoral Training in High Performance Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
+\section*{Acknowledgements}
+We thank Allen Goodman, Jane Hung, Claire McQuin, Kyle Karhohs, Beth Cimini, Tim Becker, Minh Doan, Jeanelle Ackerman and Ray Jones of the Broad Institute for fruitful discussion, guidance and support with our experimental setup.
+We also extend our gratitude to Mike Ando and Google for providing computational resources to accelerate our research.
+
+This work was supported in part by a grant from the US National Science Foundation (CAREER DBI 1148823 to AEC).
+
+Nick Pawlowski is supported by Microsoft Research through its PhD Scholarship Program and the EPSRC Centre for Doctoral Training in High Performance Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
 

### 2017-11-24 paper/mlcb16-poster/main.tex
Making poster

\documentclass[a0paper,portrait]{baposter}
\usepackage[utf8]{inputenc}

\usepackage{relsize}                  % For \smaller
\usepackage{url}  % For \url
\usepackage{tikz}  % To draw background
\usepackage{booktabs}

%%% Global Settings %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\graphicspath{{figures/}}            % Root directory of the pictures

%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Those are the official imperial colours:
\definecolor{Navy}{RGB}{0,33,71}
\definecolor{ImperialBlue}{RGB}{0,62,116}
\definecolor{LightGrey}{RGB}{235,238,238}
\definecolor{CoolGrey}{RGB}{157,157,157}
\definecolor{LightBlue}{RGB}{212,239,252}
\definecolor{Black}{RGB}{0,0,0}
\definecolor{White}{RGB}{255,255,255}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%% Utility functions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Save space in lists. Use this after the opening of the list %%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
  \setlength{\itemsep}{3pt}
  \setlength{\parskip}{2pt}
  \setlength{\parsep}{2pt}

\vspace{-0.75em}

}

\newcommand{\tab}{\hspace{0.5em}}

%%%%%% Author Affiliations %%%
\newcommand{\footremember}[2]{%
    \footnote{#2}
    \newcounter{#1}
    \setcounter{#1}{\value{footnote}}%
}
\newcommand{\footrecall}[1]{%
    \footnotemark[\value{#1}]%
}

\usepackage[backend=biber, sorting=nty, style=ieee, minbibnames=1, maxbibnames=1, maxcitenames=1, mincitenames=1, defernumbers=true]{biblatex}
\addbibresource{nips_2016.bib}
%\addbibresource{websites.bib}

\AtEveryBibitem{%
  \clearfield{url}%
  \clearfield{doi}%
  \clearfield{issn}
  \clearfield{volume}
  \clearfield{number}
  \clearfield{pages}
  \clearfield{isbn}
  \clearfield{eprint}
}

\DefineBibliographyStrings{english}{%
  references = {},
}

\renewcommand\refname{}

\begin{document}
%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\background{
  % Change colours of body background and header background if desired
  % you can also use tikz to include a custom picture here
  \begin{tikzpicture}[remember picture,overlay]%
    %the poster background colour
    \fill[fill=LightGrey] (current page.north west) rectangle (current page.south east);
    %the header
    \fill [fill=LightGrey] (current page.north west) rectangle ([yshift=-\headerheight] current page.north east);
  \end{tikzpicture}
}

\begin{poster}{
  grid=false,
  eyecatcher=true, 
  borderColor=White,
  headerColorOne=ImperialBlue,
  headerColorTwo=ImperialBlue,
  headerFontColor=LightGrey,
  boxColorOne=White,
  headershape=roundedright,
  textborder=roundedleft,
  background=user,
  headerborder=none, 
  textborder=none,
  boxshade=plain,
  headerheight=100pt
}
%%% Eye Cacther %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
  %\vspace{-20pt}
  %\includegraphics[width=0.15\textwidth]{broad}
}
%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
  {\sf\bf \textcolor{Black}{Automating Morphological Profiling with Generic Deep Convolutional Networks}}
  \vspace{3pt} % move authors list down a little bit
} 
%%% Authors %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
  \textcolor{Black}{Nick Pawlowski$^{1}$, Juan C Caicedo$^{2}$, Shantanu Singh$^{2}$, Anne E Carpenter$^{2}$, Amos Storkey$^{3}$}\\
  \vspace{3pt} \smaller[2] \textcolor{Black}{(1) Imperial College (previously (3)), (2) Broad Institute, (3) University of Edinburgh}
  %\textcolor{Black}{np716@imperial.ac.uk, \{jccaicedo, shsingh, anne\}@broadinstitute.org, a.storkey@ed.ac.uk}
}
%%% Logo %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
  %\vspace{-20pt}
  \includegraphics[width=0.2\textwidth]{logos}
}

\headerbox{1: Motivation}{name=motivation,column=0,row=0}{

Morphological profiling aims to create {\bf profiles} or signatures of genes, chemicals and diseases from microscopy images. These signatures have several applications, including
\begin{itemize}
\compresslist
\item functional genomics
\item drug discovery
\item target identification
\end{itemize}
Current approaches need human supervision to fine-tune parameters of classical algorithms to extract meaningful features from images \cite{caicedo_profiling}. We propose the use of pre-trained deep convolutional networks for feature extraction. Our contributions are:
\begin{itemize}
\compresslist
\item \textbf{Speed}: It enables faster profile computation than the classical pipeline with image segmentation and feature extraction.
\item \textbf{Autonomous}: It eliminates the need for human input to tune parameters.
\item \textbf{Performance}: The extracted profiles achieve a better accuracy than a baseline approach based on handcrafted features.
\end{itemize}
}

\headerbox{2: Overview}{name=overview, span=2, column=1, row=0}{

Activations of pre-trained neural network constitute profiles for individual images. Profiles from images of the same treatment are averaged to form a treatment (i.e. gene, chemical, or disease) profile.

\begin{minipage}[c]{\textwidth}
	\centering
	\includegraphics[width=0.95\textwidth]{overview3}
\end{minipage}
\\
\\
By measuring similarities or distances between these profiles in this representation space, they can be used to find relationships between the treatments – a valuable tool when investigating the function of a gene or allele, figuring out the mechanism of action of a compound, relating the phenotype of a disease with that of a genetic or chemical treatment, and many questions. 
}

\headerbox{3: Representation Space}{name=rep, span=1, column=0, row=1, below=motivation}{

On an example dataset, the extracted profiles from neural networks span a representation space with mostly clear clusters regarding the biological mechanism-of-action. Inception-v3 extracted profiles can be visualised using t-SNE.
\\
\\
\begin{minipage}[c]{\textwidth}
	\centering
	\includegraphics[width=1.0\textwidth]{tsne}
\end{minipage}
\\
\\
Inception-v3 profiles show the best clustering of biological mechanisms-of-action of the tested compounds.
}

\headerbox{4: Mechanism-of-Action Classification}{name=results, span=2, column=1, row=1, below=overview}{

Classification results using different pre-trained neural networks with different types of images. The neural networks were pre-trained on the ImageNet dataset. The full-sized images were resized to the appropriate input size of the neural network. Greyscale images account for the use of each greyscale channel as separate image. Illum corrected images use the technique proposed by Singh et al.\cite{Singh2014} to compensate for illumination bias during microscopy imaging. We use 1NN classification leaving out profiles from the same compound.
\\
\\
\begin{minipage}[c]{\textwidth}
   \centering
\begin{tabular}{l|rrrr}
\toprule
 & Inception-v3\cite{Szegedy2015} & ResNet-152\cite{He2015} & ResNet-101\cite{He2015} & VGG 16\cite{Simonyan2015}\\
\midrule
Full Images & $70.87\%$ &  $55.34\%$ & $57.28\%$ & $66.02\%$\\
+ illum corrected & $69.90\%$ & $65.05\%$ & $65.04\%$ & $63.11\%$ \\
+ greyscale & $86.41\%$ & $75.72\%$ & $70.87\%$ & $71.84\%$ \\
+ illum corrected \& greyscale  & \textbf{$91.26\%$} & $78.64\%$ & $79.61\%$ & $83.50\%$\\
\bottomrule
\end{tabular}
\end{minipage}
\\
\\
Confusion matrices for the Inception-v3 extracted profiles (left) show an improvement over the classical results (right)\cite{Singh2014}:
\\
\\
\begin{minipage}[c]{\textwidth}
	\centering
	\includegraphics[width=0.45\textwidth]{eval_inception_full_images_illumcorr_channelasimg.pdf}
    \hfill
    \includegraphics[width=0.45\textwidth]{mean_treatment_confusion.pdf}
\end{minipage}
\\
\\
The  different confusion matrices show that complimentary information is present in both profile types. 
}

\headerbox{5: Dose Response}{name=dose, span=2, column=1, row=2, below=results}{

The extracted profiles can be evaluated according to the dose response curves. Some compounds show chaotic relations (left). However, other compounds exhibit the expected direct relation between higher dose and stronger response (right).
\\
\begin{minipage}[c]{\textwidth}
	\centering
	\includegraphics[width=0.45\textwidth]{floxuridine_dose}
    \hfill
    \includegraphics[width=0.45\textwidth]{mitoxantrone_dose}
\end{minipage}

}

\headerbox{Acknowledgements}{name=ack, span=1, column=2, row=4, below=dose}{

We want to thank Mike Ando (Google, Inc.) for helpful discussions about the use of transfer learning for the domain of morphological profiling. We want to thank all members of the Imaging Platform for helpful discussions. AEC acknowledges NSF support (NSF CAREER DBI 1148823 to AEC).

}

\headerbox{Code \& Dataset}{name=Note, span=1, column=1, row=4, below=dose}{

Code is available at \url{github.com/carpenterlab/2016_pawlowski_mlcb}.\\
Dataset is available as BBBC021 of the Broad Bioimage Benchmark Collection at \url{data.broadinstitute.org/bbbc/BBBC021/}.\\
\vspace*{1.5em}
}

\headerbox{References}{name=References, span=1, column=0, row=4, below=rep}{
\vspace*{-1.5em}
\printbibliography
\vspace*{0.65em}
}

\end{poster}
\end{document}

### 2017-11-24 paper/paper.tex
Making poster

@@
 $$\min_G \max_D V(G, D) = \mathbb{E}_{\mathbf{x} \sim P_{\text{data}}}[\log(D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim P_{\text{noise}}}[\log(1 - D(G(\mathbf{z}))].$$
-\vspace{-1pt}
+\vspace{-2pt}
 Radford et al.
@@
 
-    \path (4.3, -0.8) coordinate [node] (f) node {$\mathbf{x}$};
-    \path (4.3, +0.8) coordinate [node] (r) node {$G(\mathbf{z})$};
+    \path (4.3, -0.8) coordinate [node] (f) node {$G(\mathbf{z})$};
+    \path (4.3, +0.8) coordinate [node] (r) node {$\mathbf{x}$};
 
@@
                        2/Convolution,%
-                       3/{\hspace{1mm}Dense(\small$\ell$)\textsuperscript{\color{Red}\textdagger}},%
-                       4/{Dense({\small$1$})}%
+                       3/{\hspace{1mm}Dense$(\ell)$\textsuperscript{\textdagger}},%
+                       4/{Dense$(1)$}%
                        } {
@@
     \draw (7, -1.5)
-          node {\footnotesize{\color{Red}\textdagger} Representations};
+          node {\footnotesize{\textdagger} Representations};
 
@@
   \caption{A simplified diagram of the modified DCGAN \cite{radford2015unsupervised} architecture we employ for our experiments.
-On a high level, the discriminator is a sequence of convolutional layers, leading to a fully connected (dense) layer with $\ell$ neurons.
+On a high level, the discriminator is a sequence of convolution layers, leading to a fully connected (dense) layer with $\ell$ neurons.
 We interpret the activations of these $\ell$ neurons as latent vectors and further feed them into a final dense layer with just one activation, forming the discriminator's output.
@@
 \section{Exploring Biological Phenotypes Using Cell Image Synthesis}\label{synthesis}
-% Exploring Biological Phenotypes Using Cell Image Synthesis
 GANs have been proven to synthesize realistic images both within and beyond the biological domain \cite{goodfellow2014generative,radford2015unsupervised,osokin2017gans}.
@@
 This property is maintained clearly in all generated images.
+
+To assess the quality of synthesized images we presented three expert biologists
+with 50 real cell images and 50 artifical cells generated with LSGAN. We
+randomized the order of images and asked the experts to judge whether each cell
+was real or fake. On average, 30\% of the time the synthetic cells were
+realistic enough to fool the human jurors into believing they were real.
 
@@
   \end{tikzpicture}
-  \caption{Examples of real BBBC021 cell images juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN.
-LSGAN synthesizes the most realistic images as judged by our expert biologists.}
+  \caption{Examples of real BBBC021 cell images juxtaposed with synthetic cell images generated with LSGAN, WGAN and DCGAN.}
   \label{fig:generated}

### 2017-11-24 paper/poster/poster.tex
Making poster

\documentclass[a0paper,portrait]{baposter}

% Packages

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{relsize}

\usepackage{enumitem}
\setlist{leftmargin=5mm}

\usepackage{tikz}
\usetikzlibrary{shapes}
\usepgflibrary{fpu}

% Graphics

\graphicspath{{../figures/}}
\definecolor{TUMBlue}{RGB}{0, 94, 184}

% Custom Commands

\newcommand{\aff}{\textsuperscript{$\star$}}
\newcommand{\compresslist}{
\setlength{\itemsep}{1pt}
\setlength{\parskip}{2pt}
\setlength{\parsep}{2pt}
\vspace{-0.75em}
}

% Bibliography

\usepackage[
  backend=biber,
  sorting=nty,
  style=ieee,
  minbibnames=1,
  maxbibnames=1,
  maxcitenames=1,
  mincitenames=1,
  defernumbers=true]{biblatex}
\addbibresource{../lit.bib}

\AtEveryBibitem{%
  \clearfield{url}%
  \clearfield{doi}%
  \clearfield{issn}
  \clearfield{volume}
  \clearfield{number}
  \clearfield{pages}
  \clearfield{isbn}
  \clearfield{eprint}
}

\DefineBibliographyStrings{english}{%
  references = {},
}

\renewcommand\refname{}

% Document

\begin{document}

% Background is white.
% \background

\background{
  \begin{tikzpicture}[remember picture,overlay]%
    \draw (12.75, 7.5) node {\includegraphics[scale=1]{tiled}};
    \draw (12.75, 24) node {\includegraphics[scale=1]{tiled}};

    \draw[white, fill=white, rounded corners=1pt]
         (current page.north west)+(1.5, -0.5)
          rectangle ++(20.5, -3.5);

    \draw[white, fill=white, rounded corners=1pt]
         (current page.north east)+(-3.2, -0.5) rectangle ++(-0.3, -3.2);

  \end{tikzpicture}
}

\begin{poster}{
  grid=false,
  eyecatcher=true,
  borderColor=white,
  headerColorOne=TUMBlue,
  headerColorTwo=TUMBlue,
  headerFontColor=white,
  boxColorOne=white,
  headershape=smallrounded,
  background=user,
  headerborder=none,
  textborder=none,
  boxshade=plain,
  headerheight=100pt
}
{
  % Eye Catcher
}
{
  {\bf CytoGAN: Generative Modeling of Cell Images}
  \vspace{3mm}
}
%%% Authors %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
  \smaller[0.5]
  \hspace{-3mm}Peter Goldsborough\textsuperscript{$\star$\,\textdagger}, Nick Pawlowski\textsuperscript{$\ddagger$}, Juan C. Caicedo\aff, Shantanu Singh\aff, Anne E. Carpenter\aff

  \vspace{2mm}
  {
    \smaller[1]
    \textsuperscript{$\star$} Broad Institute of MIT and Harvard \hspace{3mm}%
    \textdagger\hspace{2pt} Technical University Munich \hspace{3mm}%
    $\ddagger$ Imperial College London
  }
}
% Logos
{
  \begin{tabular}{l}
    \includegraphics[width=0.1\textwidth]{broad-logo}\\
    \vspace{-2mm}\\
    \includegraphics[width=0.1\textwidth]{imperial-logo}
    \vspace{-2mm}\\
    \includegraphics[width=0.1\textwidth]{tum-logo}\\
  \end{tabular}
}

\headerbox{Introduction}{name=intro, column=0, row=0}{

\emph{Morphological profiling} aims to map microscopy images of perturbed cells to salient vector representations that divide the morphological space into clusters of cells with similar properties or function.

\vspace{2mm}

Current approaches are divided between a) classical image processing, requiring manual fine-tuning and expert knowledge b) transfer learning with deep CNNs trained on miscellaneous objects, allowing no domain-specific adaptation.

\vspace{2mm}

Instead, we model cells with \emph{Generative Adversarial Networks} (GANs) for both representation learning and cell synthesis. Advantages of our approach are:
\vspace{2mm}
\begin{itemize}
\compresslist
\item \textbf{Unsupervised} representation learning, which can readily incorporate new data;
\item \textbf{Specialization} to training data, allowing identification of the semantic relations between biologically meaningful channels;
\item \textbf{Interpretability} of learned representations and useful visualization that help translate data variations into meaningful biological phenotypes.
\end{itemize}
}

\headerbox{Exploring Biological Phenotypes Using Cell Synthesis}{name=synth, span=2, column=1, row=0}{

Besides learning salient representations of microscopy cell images, GANs allow us to synthesize cell images and explore variations in the noise and learned latent spaces. This allows biologically meaningful experiments and improves interpretability of our method.

\vspace{-1mm}
\begin{center}
  \begin{tikzpicture}
      \foreach \i/\kind in {0/began, 1/bigan, 2/lsgan, 3/real, 4/wgan} {
        \foreach \j in {0} {
          \draw ({\i * 3}, {\j * 1.5})
          node {\includegraphics[width=2cm, height=2cm]{generated/\kind/\j}};
        }
      }
  \end{tikzpicture}
\end{center}
\vspace{-2mm}

Shown above are cells synthesized with LSGAN, WGAN, BEGAN and BiGAN
architectures alongside a real cell. The synthesized images are not only highly
detailed and realistic, but also consistent with their biological nature. For
example, it is characteristic that $\beta$-Tubulin (green channel) forms a
circular halo cradling the nucleus. This property is maintained clearly in all
generated images. }

\headerbox{Interpolating Cell Images in Noise and Latent Space}{name=interpol, span=2, column=1, row=1, below=synth}{

GANs are known to learn low manifolds of their input priors, such that
interpolation between two points $\mathbf{z}_1, \mathbf{z}_2$ drawn from the noise space $P_{\text{noise}}$ results in visually smooth
transitions in generated images. We confirm that this property holds for
microscopy cell images.

\vspace{-1mm}
\begin{center}
\begin{tikzpicture}
  \foreach \i in {0, 1} {
    \foreach \j in {0, ..., 9} {
      \draw ({\j * 1.5}, {\i * -1.5}) node
      {\includegraphics[width=1.5cm, height=1.5cm]{noise-interpolation/\i/\j}};
    }

  }
\end{tikzpicture}
\end{center}
\vspace{-2mm}

Bidirectional GANs (BiGAN\cite{}) extend the vanilla DCGAN architecture with an
explicit encoder network that enables both synthesis of images from noise
\emph{and} inference of latent representations, in such a way that the noise and
latent space are fused. This means we can encode real cell images, perturb,
transform or, in this case, interpolate between them, and re-synthesize the
resulting vectors.

\vspace{-2mm}
\begin{center}
\begin{tikzpicture}
    \foreach \i in {0, ..., 9} {
      \draw ({\i * 1.5}, 0) node
      {\includegraphics[width=1.5cm, height=1.5cm]{latent-interpolation/\i}};
    }

    \fill[black] (0.25, -0.75) rectangle ++(12, 2mm);
\end{tikzpicture}
\end{center}
}

\headerbox{Vector Algebra for Biological Interpretability}{name=algebra, span=2, column=1, row=1, below=interpol}{

We find that GANs partition the noise space into regions capturing different semantic properties of cells, allowing biologically meaningful algebra on noise vectors.

\vspace{-3mm}
\begin{center}
\begin{tikzpicture}
    \draw (0, 0) node {\includegraphics[scale=0.4]{algebra/figure-chan}};
    \draw (-3.7, -1.2) node {\footnotesize Much $\beta$-Tubulin}
        ++(2.5, 0) node {\footnotesize Little $\beta$-Tubulin}
        ++(2.5, 0) node {\footnotesize Any Cell}
        ++(2.5, 0) node {\footnotesize More $\beta$-Tubulin}
        ;

    % \draw (0, -2.5) node {\includegraphics[scale=0.4]{algebra/figure-nucl}};
    % \draw (-3.7, -3.7) node {\footnotesize Large Nucleus}
    %     ++(2.5, 0) node {\footnotesize Small Nucleus}
    %     ++(2.5, 0) node {\footnotesize Any Cell}
    %     ++(2.5, 0) node {\footnotesize Larger Nucleus}
        ;
\end{tikzpicture}
\end{center}
\vspace{-3mm}

When the noise and learned latent space are fused (BiGAN), algebra on real cell images becomes possible. This open the door to discovery of a whole new class of highly interpretable and biologically valuable relationships, such as:

\begin{tikzpicture}
  \draw (-8, 0);
  \draw (0, 0) node [text width=10cm] {
  \begin{itemize}
    \compresslist
    \item $\text{emetine}_{1.0} - \text{emetine}_{0.3} + \text{taxol}_{0.3} \stackrel{?}{=} \text{taxol}_{1.0}$
    \item $\text{Protein Synthesis} - \text{Protein Degradation} \stackrel{?}{=} \text{DMSO}$
    \item $\text{Kinase Inhibitor} - \text{DMSO} + \text{DMSO}' \stackrel{?}{=} \text{Kinase Inhibitor}$
  \end{itemize}
  };
\end{tikzpicture}
\vspace{2.5pt}
}

\headerbox{Representation Learning for Morphological Profiling}{name=rep, span=2, column=1, row=1, below=algebra}{

We investigate the ability of GANs to learn vector representations of cell
images and evaluate their quality quantitatively at the task of
mechanism-of-action (MOA) prediction. We train a variety of GAN models on 1.3
million microscopy images of cells perturbed with particular drug compounds,
average signatures across treatments and predict the MOA via nearest-neighbor
classification.

\vspace{2mm}
\begin{tikzpicture}
  \draw (-2, 0);
  \draw (0, 0.3) node {\includegraphics[scale=0.2, trim={4cm 3.5cm 8cm 3cm}, clip]{latent-space}};
  \draw (7.75, 1.2) node [text width=11cm, align=justify] {
  For WGAN and LSGAN we interpret the activations of the penultimate dense
  layer as a source of representations. BEGAN an BiGAN architectures already have explicit inference modules. We find that while GANs
  are superior to autoencoders, they are not yet competitive with existing
  classical or deep transfer-learning techniques.
  };

  \draw (8, -1) node [text width=10.5cm] {
    \begin{tabular}{ccccc}
      \toprule
      LSGAN & BiGAN & VAE \cite{pawlowski2016msc} & CP \cite{Singh2014} & Transfer Learning \cite{ando2017improving} \\
      \midrule
      68\% & 70\%& 49\% & 90\% & 96\% \\
      \bottomrule \\
    \end{tabular}
  };
\end{tikzpicture}
\vspace{-6mm}
}

\headerbox{Code \& Data}{name=code, column=0, row=1, below=intro}{Code and experiments are available at \url{github.com/goldsborough/cytogan}. Our data is the BBBC021 dataset from the Broad BioImage Benchmark Collection (\url{data.broadinstitute.org/bbbc/BBBC021/}).}

\headerbox{Acknowledgements}{name=ack, column=0, row=2, below=code}{
We thank Mike Ando and Google, Inc. for providing computational resources to accelerate our research. We also thank all members of the Broad Imaging Platform for fruitful discussion and their continuous guidance and support.
}
\headerbox{References}{name=ref, column=0, row=3, below=ack}{
\printbibliography
}

\end{poster}
\end{document}

### 2017-11-28 paper/paper.tex
Updates to paper and poster

@@
   Cambridge, MA, USA \\
-  \texttt{\{jccaicedo, shsingh, anne\}@broadinstitute.org} \\
+  \texttt{\{jcaicedo, shsingh, anne\}@broadinstitute.org} \\
 }
@@
                        2/Convolution,%
-                       3/{\hspace{1mm}Dense$(\ell)$\textsuperscript{\textdagger}},%
+                       3/{\hspace{1mm}Dense$(\ell)$\textsuperscript{$\star$}},%
                        4/{Dense$(1)$}%
@@
     \draw (7, -1.5)
-          node {\footnotesize{\textdagger} Representations};
+          node {\footnotesize{\textsuperscript{$\star$}}\hspace{-1pt} Representations};
 
@@
 characteristic that $\beta$-Tubulin forms a circular halo cradling the nucleus.
-This property is maintained clearly in all generated images.
+This property is maintained clearly in most generated images.
 
@@
 We believe that improved inference, combined with the interpretable nature of the GAN framework, may enable simulated experiments by performing algebra with vectors corresponding to cell lines, diseases or other perturbations.
-Finally, we note that deep learning generally flourishes with larger amounts of data.
-Thus, we conjecture that the performance of the outlined approaches will improve with larger datasets.
 

### 2017-11-28 paper/poster/poster.tex
Updates to paper and poster

@@
 \usepackage[utf8]{inputenc}
+\usepackage[american]{babel}
+
 \usepackage{url}
@@
   mincitenames=1,
-  defernumbers=true]{biblatex}
+  defernumbers=true,
+  natbib=true]{biblatex}
 \addbibresource{../lit.bib}
@@
   \begin{tikzpicture}[remember picture,overlay]%
-    \draw (12.75, 7.5) node {\includegraphics[scale=1]{tiled}};
-    \draw (12.75, 24) node {\includegraphics[scale=1]{tiled}};
+    % \draw (12.75, 7.5) node {\includegraphics[scale=1]{tiled}};
+    % \draw (12.75, 24) node {\includegraphics[scale=1]{tiled}};
 
@@
     \draw[white, fill=white, rounded corners=1pt]
-         (current page.north east)+(-3.2, -0.5) rectangle ++(-0.3, -3.2);
+         (current page.north east)+(-3.3, -0.5) rectangle ++(-0.4, -3.2);
 
@@
     \smaller[1]
-    \textsuperscript{$\star$} Broad Institute of MIT and Harvard \hspace{3mm}%
+    \aff Broad Institute of MIT and Harvard \hspace{3mm}%
     \textdagger\hspace{2pt} Technical University Munich \hspace{3mm}%
@@
   \begin{tabular}{l}
-    \includegraphics[width=0.1\textwidth]{broad-logo}\\
+    \includegraphics[width=0.11\textwidth]{broad-logo}\\
     \vspace{-2mm}\\
-    \includegraphics[width=0.1\textwidth]{imperial-logo}
+    \includegraphics[width=0.11\textwidth]{imperial-logo}
     \vspace{-2mm}\\
-    \includegraphics[width=0.1\textwidth]{tum-logo}\\
+    \includegraphics[width=0.11\textwidth]{tum-logo}\\
   \end{tabular}
@@
 
-Current approaches are divided between a) classical image processing, requiring manual fine-tuning and expert knowledge b) transfer learning with deep CNNs trained on miscellaneous objects, allowing no domain-specific adaptation.
+Current approaches are divided between \textbf{a}) classical image processing, requiring manual fine-tuning and expert knowledge, and \textbf{b}) transfer learning with deep CNNs trained on miscellaneous objects, allowing little to no domain-specific adaptation.
 
@@
 
-Instead, we model cells with \emph{Generative Adversarial Networks} (GANs) for both representation learning and cell synthesis. Advantages of our approach are:
+Instead, we model cells with \emph{Generative Adversarial Networks} (GANs) for both representation learning and cell image synthesis. Advantages of our approach include:
 \vspace{2mm}
@@
 example, it is characteristic that $\beta$-Tubulin (green channel) forms a
-circular halo cradling the nucleus. This property is maintained clearly in all
+circular halo cradling the nucleus. This property is maintained clearly in most
 generated images. }
@@
 interpolation between two points $\mathbf{z}_1, \mathbf{z}_2$ drawn from the noise space $P_{\text{noise}}$ results in visually smooth
-transitions in generated images. We confirm that this property holds for
-microscopy cell images.
+transitions in generated images \cite{radford2015unsupervised}. We confirm that this property holds for microscopy cell images.
 
@@
 
-Bidirectional GANs (BiGAN\cite{}) extend the vanilla DCGAN architecture with an
+Bidirectional GANs (BiGANs) \cite{donahue2016adversarial} extend the vanilla DCGAN \cite{radford2015unsupervised} architecture with an
 explicit encoder network that enables both synthesis of images from noise
@@
 latent space are fused. This means we can encode real cell images, perturb,
-transform or, in this case, interpolate between them, and re-synthesize the
-resulting vectors.
+transform or in this case interpolate between them and re-synthesize the
+resulting images.
 
@@
 
-When the noise and learned latent space are fused (BiGAN), algebra on real cell images becomes possible.
-This open the door to discovery of a whole new class of highly interpretable and biologically valuable relationships, such as:
+When the noise and learned latent space are fused (BiGAN), algebra on real cell
+images becomes possible. This opens the door to discovery of a whole new class
+of highly interpretable and biologically valuable relationships, such as:
 
+\vspace{1mm}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
-\vspace{2.5pt}
+\vspace{-2mm}
 }
-
 \headerbox{Representation Learning for Morphological Profiling}{name=rep, span=2, column=1, row=1, below=algebra}{
 
-We investigate the ability of GANs to learn vector representations of cell
-images and evaluate their quality quantitatively at the task of
+We investigate the ability of GANs to learn high quality vector representations
+of cell images and evaluate them quantitatively at the task of
 mechanism-of-action (MOA) prediction. We train a variety of GAN models on 1.3
@@
   \draw (0, 0.3) node {\includegraphics[scale=0.2, trim={4cm 3.5cm 8cm 3cm}, clip]{latent-space}};
-  \draw (7.75, 1.2) node [text width=11cm, align=justify] {
-  For WGAN and LSGAN we interpret the activations of the penultimate dense
-  layer as a source of representations. BEGAN an BiGAN architectures already have explicit inference modules. We find that while GANs
-  are superior to autoencoders, they are not yet competitive with existing
-  classical or deep transfer-learning techniques.
-  };
+  % \draw (7.75, 1.2) node [text width=11cm, align=justify] {
+  % For LSGAN we interpret the activations of the penultimate dense
+  % layer as a source of representations. BiGAN has an explicit inference module and requires no further modification. We find that while GANs
+  % are superior to autoencoders, they are not yet competitive with existing
+  % classical or deep transfer-learning techniques at this task.
+  % };
+
+  \draw (7.75, 1.2) node {\scalebox{0.8}{
+    \begin{tikzpicture}[thick]
+      \tikzset{box/.style={
+        draw, rectangle, rounded corners=0.8pt,
+        text width=2.2cm, text height=0.75cm}}
+      \tikzset{node/.style={draw, circle, inner sep=0.35cm}}
+
+      % Spacer
+      \draw (10.5, 0);
+
+      % Model
+      \path (0, 0) coordinate [draw, circle, inner sep=5pt]
+            (z) node {$\mathbf{z}$};
+
+      \path (2, 0) coordinate [box] (G) node {Generator};
+
+      \path (4.3, -0.8) coordinate [node] (f) node {$G(\mathbf{z})$};
+      \path (4.3, +0.8) coordinate [node] (r) node {$\mathbf{x}$};
+
+      \path (9.5, 0) coordinate [draw, circle, inner sep=7pt]
+            (D) node {$\hspace{1pt}\mathbf{D}$};
+
+      % Anchor for real and fake arrows.
+      \path (5, 0) coordinate (_);
+
+      \foreach \x/\l in {0/Convolution,%
+                         1/Convolution,%
+                         2/Convolution,%
+                         3/{\hspace{1mm}Dense$(\ell)$\textsuperscript{$\star$}},%
+                         4/{Dense$(1)$}%
+                         } {
+        \path ({5.8 + \x * 0.65}, 0)
+              coordinate [box, text width=0.5cm, text height=2cm] (l\x)
+              node [rotate=90] {\l};
+      }
+
+      % Footnote
+      \draw (7, -1.5)
+            node {\footnotesize{\textsuperscript{$\star$}}\hspace{-1pt} Representations};
+
+      % Edges
+      \draw (f) -- (_);
+      \draw (r) -- (_);
+      \draw [->] (_) -- (l0);
+      \draw [->] (l4) -- (D);
+      \draw [->] (z) -- (G);
+      \draw [->] (G) edge[out=0] (f);
+    \end{tikzpicture}
+  }};
 
@@
 
-\headerbox{Code \& Data}{name=code, column=0, row=1, below=intro}{Code and experiments are available at \url{github.com/goldsborough/cytogan}.
-Our data is the BBBC021 dataset from the Broad BioImage Benchmark Collection (\url{data.broadinstitute.org/bbbc/BBBC021/}).}
+\headerbox{Code \& Data}{name=code, column=0, row=1, below=intro}{
+Code and experiments are available at \url{github.com/goldsborough/cytogan}. Our data is the BBBC021 dataset from the Broad BioImage Benchmark Collection (\url{data.broadinstitute.org/bbbc/BBBC021}).
+
+% REPLACE WITH PAPER!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+\vspace{-3mm}
+\begin{center}
+  \includegraphics[scale=0.174]{qr-code}\hspace{1cm}\includegraphics[scale=0.174]{qr-code}
+
+  \vspace{-3mm}
+  \hspace{0.1cm}\texttt{Code}\hspace{2.6cm}\texttt{Paper}
+\end{center}
+
+% \includegraphics[scale=0.1]{qr-paper}
+}
 
@@
 }
+
 \headerbox{References}{name=ref, column=0, row=3, below=ack}{
+\vspace{-5mm}
 \printbibliography

### 2017-12-01 paper/poster/poster.tex
Small update to paper and poster
Fixed poster

@@
 \compresslist
-\item \textbf{Unsupervised} representation learning, which can readily incorporate new data;
-\item \textbf{Specialization} to training data, allowing identification of the semantic relations between biologically meaningful channels;
-\item \textbf{Interpretability} of learned representations and useful visualization that help translate data variations into meaningful biological phenotypes.
+\item \textbf{Adaptability} to training data, allowing identification of the intrinsic semantic relationships between biologically meaningful channels,
+\item \textbf{Translation} of learned representations into biological phenotypes.
 \end{itemize}
@@
 
-Besides learning salient representations of microscopy cell images, GANs allow us to synthesize cell images and explore variations in the noise and learned latent spaces.
-This allows biologically meaningful experiments and improves interpretability of our method.
+GANs enable us to synthesize cell images and explore variations in the noise and
+learned latent spaces. This allows biologically meaningful experiments and
+improves interpretability of our method.
 
@@
     \item $\text{Protein Synthesis} - \text{Protein Degradation} \stackrel{?}{=} \text{DMSO}$
-    \item $\text{Kinase Inhibitor} - \text{DMSO} + \text{DMSO}' \stackrel{?}{=} \text{Kinase Inhibitor}$
+    \item $\text{Kinase Inhibitor} - \text{DMSO} + \text{DMSO}' \stackrel{?}{=} \text{Kinase Inhibitor}'$
   \end{itemize}
@@
 
+% We investigate the ability of GANs to learn high quality vector representations
+% of cell images and evaluate them quantitatively at the task of
+% mechanism-of-action (MOA) prediction. We train a variety of GAN models on 1.3
+% million microscopy images of cells perturbed with particular drug compounds,
+% average signatures across treatments and predict the MOA via nearest-neighbor
+% classification.
+
 We investigate the ability of GANs to learn high quality vector representations
-of cell images and evaluate them quantitatively at the task of
-mechanism-of-action (MOA) prediction. We train a variety of GAN models on 1.3
-million microscopy images of cells perturbed with particular drug compounds,
-average signatures across treatments and predict the MOA via nearest-neighbor
-classification.
+of cells perturbed with particular drug compounds. We want chemicals with
+similar effects on cells to have short distances in feature space. We train a variety of
+GAN models on 1.3 million cell images, average signatures across treatments, and
+predict the effect of unseen chemicals via nearest-neighbor classification.
 
-\vspace{2mm}
+\vspace{-2mm}
 \begin{tikzpicture}
   \draw (-2, 0);
-  \draw (0, 0.3) node {\includegraphics[scale=0.2, trim={4cm 3.5cm 8cm 3cm}, clip]{latent-space}};
-  % \draw (7.75, 1.2) node [text width=11cm, align=justify] {
-  % For LSGAN we interpret the activations of the penultimate dense
-  % layer as a source of representations. BiGAN has an explicit inference module and requires no further modification. We find that while GANs
-  % are superior to autoencoders, they are not yet competitive with existing
-  % classical or deep transfer-learning techniques at this task.
-  % };
+  \draw (0, 0.6) node {\includegraphics[scale=0.2, trim={4cm 3.5cm 8cm 3cm}, clip]{latent-space}};
 
-  \draw (7.75, 1.2) node {\scalebox{0.8}{
+  \draw (7.75, -0.5) node {\scalebox{0.8}{
     \begin{tikzpicture}[thick]
@@
       % Footnote
-      \draw (7, -1.5)
-            node {\footnotesize{\textsuperscript{$\star$}}\hspace{-1pt} Representations};
+      \draw (1.9, -1)
+            node {\small{\textsuperscript{$\star$}}\hspace{-1pt} Representations};
 
@@
 
-  \draw (8, -1) node [text width=10.5cm] {
+  \draw (8, 2) node [text width=10.5cm] {
+    \begin{center}\textbf{Prediction Accuracy}\end{center}
+      \vspace{-1mm}
     \begin{tabular}{ccccc}
@@
       \midrule
-      68\% & 70\%& 49\% & 90\% & 96\% \\
+      68\%/76\%\textsuperscript{\textdagger} & 70\%/72\%\textsuperscript{\textdagger} & 49\% & 90\% & 91\%/96\%\textsuperscript{\textdagger} \\
       \bottomrule \\
@@
   };
+
+  \draw (4.4, 0.7) node {\footnotesize\textdagger\hspace{0.5pt} Whitening Transform};
 \end{tikzpicture}
@@
 % REPLACE WITH PAPER!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-\vspace{-3mm}
-\begin{center}
-  \includegraphics[scale=0.174]{qr-code}\hspace{1cm}\includegraphics[scale=0.174]{qr-code}
-
-  \vspace{-3mm}
-  \hspace{0.1cm}\texttt{Code}\hspace{2.6cm}\texttt{Paper}
-\end{center}
-
-% \includegraphics[scale=0.1]{qr-paper}
+% \vspace{-3mm}
+% \begin{center}
+%   \includegraphics[scale=0.08]{qr-code}\hspace{1cm}\includegraphics[scale=0.08]{qr-code}
+%
+%   \vspace{-3mm}
+%   \hspace{0.1cm}\texttt{Code}\hspace{2.6cm}\texttt{Paper}
+% \end{center}
 }
@@
 We also thank all members of the Broad Imaging Platform for fruitful discussion and their continuous guidance and support.
+
+\vspace{2mm}
+
+This work was supported in part by a grant from the US National Science Foundation (CAREER DBI 1148823 to AEC).
+
+\vspace{1mm}
+
+Nick Pawlowski is supported by Microsoft Research through its PhD Scholarship Program and the EPSRC Centre for Doctoral Training in High Performance Embedded and Distributed Systems (HiPEDS, Grant Reference EP/L016796/1).
 }

### 2017-12-01 paper/paper.tex
Small update to paper and poster
Fixed a few things in the paper

@@
 Our work focuses on the task of morphological profiling, which aims to map microscopy images of cells to representation vectors that capture salient features and axes of variation in an unsupervised manner \cite{caicedo_profiling}.
-These representations ideally divide the morphological space into clusters of cells with similar properties or, in the case of induced perturbations, similar function.
+These representations ideally divide the morphological space into clusters of cells with similar properties or, in the case of chemical or genetic treatments added to the cells, similar function.
 
@@
 \begin{itemize}
-  \item \emph{Adaptability}: Unsupervised representation learning, which can readily incorporate new data.
-This is important given that applications of morphological profiling usually lack ground truth annotations, but instead aim to reveal the structure of the image collection.
-  \item \emph{Specialization}: Because this method learns from its training data it is able to disentangle factors of variation and extract inherent semantic relationships.
+  \item \emph{Adaptability}: Because this method learns from its training data it is able to disentangle factors of variation and extract inherent semantic relationships.
 Transfer learning approaches lack this ability and cannot capture the intrinsic relations between the biologically meaningful channels.
@@
 We stack these channels and treat them as RGB images via a simple $\text{DNA} \mapsto \text{R}, \text{$\beta$-Tubulin} \mapsto \text{G}, \text{F-Actin} \mapsto \text{B}$ mapping.
-We obtain ca.
-1.3 million single cell images by segmenting the raw images using CellProfiler.
+We obtain 1.3 million single cell images by segmenting the raw images using CellProfiler.
 Finally, we normalize the luminance of each channel to the range [0 -- 1].
@@
 realistic, but also consistent with their biological nature. For example, it is
-characteristic that $\beta$-Tubulin forms a circular halo cradling the nucleus.
+characteristic that $\beta$-Tubulin (green) forms a circular halo cradling the nucleus.
 This property is maintained clearly in most generated images.
@@
 To assess the quality of synthesized images we presented three expert biologists
-with 50 real cell images and 50 artifical cells generated with LSGAN. We
+with 50 real cell images and 50 artificial cells generated with LSGAN. We
 randomized the order of images and asked the experts to judge whether each cell
